---
title: "data_clustering"
output: html_document
date: "2026-04-25"
params:
  use_parent: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(code_folding = "show")
knitr::opts_knit$set(root.dir = '~/PycharmProjects/corpus-caesarianum-authorship')
```

```{r Libraries}


library(factoextra)
library(ggrepel)
library(tidyverse)
library(cluster)
library(ggplot2)
library(magrittr)
library(combinat)
library(dendextend)
library(MASS)
library(dplyr)
library(multcomp)
library(ggtext)
library(printr)
library(ggdendro)
library(stringr)
library(logger)
```

## Get the data

This data was created with the data-processing.Rmd notebook, which arranged the texts for clustering. This means that each row is an observation (a text), and each column a variable. `simplex_vars_only` generates a vector of only length 865 (as each variable is just the count of each individual feature, such as accusative nouns, participles, 3rd person verbs, and so on). The gorman.vars variables consider the top 1,000 combinations of features, so a single feature might be "masculine, ablative, oblique argument with a 3rd person verb parent".

```{r}

simplex_vars_only <- read.csv("./feature_csv_files/feature_values.csv")

file.vars.gorman <- if (params$use_parent) "./feature_csv_files/feature_values_gorman.csv" else "./feature_csv_files/feature_values_wordonly.csv"

gorman.vars <- read.csv(file.vars.gorman); gorman.vars$title <- gsub(" ", "", gorman.vars$title)

file.sectioned.gorman <- if (params$use_parent) "./feature_csv_files/feature_values_gorman_by_section.csv" else "./feature_csv_files/feature_values_wordonly_by_section.csv"

sectioned.gorman.vars <- read.csv("./feature_csv_files/feature_values_gorman_by_section.csv");sectioned.gorman.vars$title <- gsub(" ", "", sectioned.gorman.vars$title)

# add_book <- function(source_data) {
#   source_data |>
#     mutate(book = paste0((strsplit(title, "_"))[[1]][1:2], collapse = "_")) |>
#     relocate(book, .before = 2)
# }

# sectioned.gorman.vars <- add_book(sectioned.gorman.vars)

cicero_works <- "(philippics|senectute|amicitia)"

sectioned.gorman.vars <- bind_rows(
  filter(sectioned.gorman.vars, str_starts(title, cicero_works)),
  filter(sectioned.gorman.vars, !str_starts(title, cicero_works))
  )

gorman.vars <- bind_rows(
  filter(gorman.vars, str_starts(title, cicero_works)),
  filter(gorman.vars, !str_starts(title, cicero_works))
)

#cic.gorman.vars <- read.csv("./feature_csv_files/feature_values_cicero.csv")K-d
```

```{r}
#For converting to readable names in graph labels.
lookup_table <- setNames(
  c("BAlex. 1-21", "BAlex. 22-78", "BAfr.", "De Amicitia", "BC 1", "BC 2", "BC 3", "BG 1", "BG 2", "BG 3", "BG 4", "BG 5", "BG 6", "BG 7", "BG 8", "Philippics 1", "Philippics 2", "Philippics 3", "Philippics 4", "Philippics 5", "Philippics 6", "Philippics 7", "Philippics 8", "Philippics 9", "Philippics 10", "Philippics 11", "Philippics 12", "Philippics 13", "Philippics 14", "De Senectute", "BHisp."),
  c("alexandrine_1", "alexandrine_2", "african_1", "amicitia_1", "civil_1", "civil_2", "civil_3", "gallic_1", "gallic_2", "gallic_3", "gallic_4", "gallic_5", "gallic_6", "gallic_7", "gallic_8", "philippics_1", "philippics_2", "philippics_3", "philippics_4", "philippics_5", "philippics_6", "philippics_7", "philippics_8", "philippics_9", "philippics_10", "philippics_11", "philippics_12", "philippics_13", "philippics_14", "senectute_1", "spanish_1")
)
```

What about euclidean or manhattan distance?

Here's a transposed version for correlation coefficient calculation.

```{r}

t.sectioned.gorman.vars <- sectioned.gorman.vars |> 
  as.data.frame() |>
  dplyr::select(-title) |>
  t()

colnames(t.sectioned.gorman.vars) <- sectioned.gorman.vars$title

t.sectioned.gorman.vars <- t.sectioned.gorman.vars |> as.data.frame()


```

Now that we have a viable DF, let's cut it up a little bit. We don't want the first column, because that's all text. Not every clustering algorithm needs the DF to be pre-scaled (that is, subtracting the mean from every value and dividing it by the standard deviation to make the comparable). However, some do expect the data to be scaled, so lets create two variables. **df_clust** isn't scaled, and **df_scaled** is.

```{r df_clust & df_scaled}
#SELECTED ARBITRARY SUM!!!
simplex_vars_only <- slice_sample(simplex_vars_only, n = 100)
df_clust <- simplex_vars_only |> dplyr::select(-title) |> dplyr::select(where(~ sum(.x) > 5))
gorman_clust <- gorman.vars |> dplyr::select(-title)
rownames(df_clust) <- simplex_vars_only$title
# df_clust <- slice_sample(df_clust, n=45)

df_scaled <- as.data.frame(scale(df_clust))
```

# Analysis

First, to get some of the formatting predefined.

```{r}


# Add this with the + operator to fviz_cluster plots. @param data is the original dataset used for cluster analysis (df_clust in the first case)
cluster_labels <- geom_text_repel(aes(label = rownames(data)), vjust = "inward", hjust = "inward", max.overlaps = 20, size = 3)
# Args for fviz_clust
palette <- c("#2E9FDF", "#00AFBB", "#E7B800", "red", "gray", "cyan", "black")
geom <- "point"
ellipse.type <- "convex"
ggtheme <- theme_bw()

# K-means
k <- 3

# Hcut
hc_func <- "agnes"
hc_method <- "ward"
hc_metric <- "euclidean"
cex <- 0.2

# Text labels for distance diagrams. @param label is usually the rownames of a data.frame
margin_size <- 10
distance_labels <- theme(axis.text = element_text(size = 5, vjust = 1), margins = margin(t = margin_size, r = margin_size, b = margin_size, l = margin_size))

# Dendrogram theme
dend_size <- 0.2
dend_theme <- theme(axis.text = element_text(size = dend_size, vjust = 1), margins = margin(t = margin_size, r = margin_size, b = margin_size, l = margin_size))
```

## Procedure for Evaluating Clustering Methods

*copy over from Obsidian notebook*

## K-means clustering

Time to cluster! K-means is a good starting point here. I'll check out CLARA as well. We'll need to evaluate what's best. Here's the modifications we can make:

1.  Change which variables we account for.

2.  Change the number of centers (between 3 and 5? But we already have a way of visualizing that)

3.  Change the measuring method (i.e. Euclidean, Manhattan, or even cosine and correlation coefficient-based measures)

4.  Change the algorithm (use agglomerative hierarchical clustering as another method of visualizing the data)

```{r K-means clustering-single-variables}
set.seed(126126)
kmeans_basic_cluster <- kmeans(df_clust, centers = k, nstart = 50)
kmeans_basic_table <- data.frame(kmeans_basic_cluster$size, kmeans_basic_cluster$centers)
kmeans_basic_df <- data.frame(Clusters = kmeans_basic_cluster$cluster, df_clust)

labs_kmeans <- lookup_table[final.frame$title]
```

Let's do our first visualization. Four clusters seems like a reasonable amount, looking at the fviz_nbclust results with the wss method.

I would note that the inclusion or exclusion of Cicero has something of an effect on the pca. This is to be expected, and is in some ways a desirable characteristic, as we included it to make sure there was another point of comparison to show how coherent Caesar really looks when compared to an outside text. However, what do the results look like without Cicero? I need to record the data both ways.

```{r}
# TODO ADD SUMMARIES OF THE ABOVE DATA HERE

# Principal component analysis

pca <- prcomp(df_clust)

vars <- dimnames(pca$rotation)[[1]]

# View(as_tibble(pca$rotation) |> mutate(names = vars) |> relocate(names, .before = 1) |> slice_max(order_by = PC1, n = 30))

summary(pca)

# Visualizations

fviz_nbclust(df_clust, kmeans, method = "wss", nstart = 50)

dim1 <- as.character(summary(pca)$importance[2] * 100)
dim2 <- as.character(summary(pca)$importance[5] * 100)

cluster_main <- fviz_cluster(kmeans_basic_cluster,
  data = df_clust,
  palette = palette,
  geom = geom,
  ellipse.type = ellipse.type,
  ggtheme = ggtheme
) +
  geom_text_repel(
    aes(label = labs_kmeans),
    vjust = "inward",
    hjust = "inward",
    max.overlaps = 14,
    size = 2,
    seed = 12927
  ) +
  labs(
    x = paste(strwrap(paste("Dim1", dim1, "(Left = more 1st/2nd person, imperfective verbs, Right = more Ablative nouns, perfective verbs, and nouns overall)"), 70),
      collapse = "\n"
    ),
    y = paste(strwrap(paste("Dim2", dim2, "(Up = more 3rd person and past tense verbs, Down = more nouns)"), 50), collapse = "\n"),
    title = "(1) K-means clustering applied to PHI Latin Texts, k=4"
  ) +
  theme(axis.title = element_text(size = 12))

cluster_main


ggsave(filename = "Images/cluster-diagram-kmeans.png", dpi = "print")

str(kmeans_basic_cluster)

```

```{r}
fviz_dend(hclust(dist(df_clust)),
  palette = palette,
  geom = geom,
  ellipse.type = ellipse.type,
  ggtheme = ggtheme,
  cex = 0.4,
) 

ggsave(filename = "Images/dendrogram.png", dpi = "print")
```

## Exploratory Visualization

TO DO

## Gorman-Style Variables

```{r K-means clustering-combined-variables}
set.seed(126126)

gorman_clust$title <- gorman.vars$title
gorman_clust <- gorman_clust[,c(title, unlist(top_colnames))]

#Uncomment this when working with section numbers
#gorman_clust <- gorman_clust |> filter(str_detect(title, "(spanish |gallic _ 1)"))

labs_kmeans <- lookup_table[gorman_clust$title]

gorman_clust <- gorman_clust |> dplyr::select(-title)
dist_gorman <- get_dist(gorman_clust, method = "spearman")

kmeans_basic_cluster <- pam(dist_gorman, k = k, diss = TRUE)
kmeans_basic_cluster$data <- gorman_clust

#kmeans_basic_cluster <- kmeans(gorman_clust, centers = k, nstart = 25)
kmeans_basic_table <- data.frame(kmeans_basic_cluster$size, kmeans_basic_cluster$centers)
kmeans_basic_df <- data.frame(Clusters = kmeans_basic_cluster$cluster, gorman_clust)
```

Let's do our first visualization. Four clusters seems like a reasonable amount, looking at the fviz_nbclust results with the wss method.

```{r}
# TODO ADD SUMMARIES OF THE ABOVE DATA HERE

# Principal component analysis

pca <- prcomp(gorman_clust)

# I'm adding the following code after this cell was created.
# I want to test what the clustering looks like if we cull to only the top variables in pca

vars <- dimnames(pca$rotation)[[1]]

top_vars <- as_tibble(pca$rotation) |> mutate(names = vars) |> relocate(names, .before = 1) |> slice_max(order_by = PC1, n = 30)

top_colnames <- top_vars$names[1:5]
  
summary(pca)



# Visualizations

fviz_nbclust(gorman_clust, kmeans, method = "wss", nstart = 50)
fviz_nbclust(gorman_clust, kmeans, method = "silhouette", nstart = 50)
#fviz_nbclust(gorman_clust, kmeans, method = "gap", nstart = 50)

dim1 <- as.character(summary(pca)$importance[2] * 100)
dim2 <- as.character(summary(pca)$importance[5] * 100)

cluster_main <- fviz_cluster(kmeans_basic_cluster,
  data = gorman_clust,
  palette = palette,
  geom = geom,
  ellipse.type = ellipse.type,
  ggtheme = ggtheme
) +
  geom_text_repel(
    aes(label = labs_kmeans),
    vjust = "inward",
    hjust = "inward",
    max.overlaps = 14,
    size = 4,
    seed = 12927
  ) +
  labs(
    x = paste(strwrap(paste("Dim1", dim1, ""), 70),
      collapse = "\n"
    ),
    y = paste(strwrap(paste("Dim2", dim2, ""), 50), collapse = "\n"),
    title = "K-means clustering applied to the Corpus Caesarianum and Cicero, k= 3"
  ) +
  theme(axis.title = element_text(size = 12))

cluster_main


ggsave(filename = "Images/cluster-diagram-gorman-style.png", dpi = "print")

str(kmeans_basic_cluster)

```

Datanovia suggests eclust as a universal function for clustering, which performs some of the above operations together:

```{r}

eclust(x = gorman_clust, "kmeans", hc_metric = "euclidean")
```

The rotations suggest that the explanation of the data is poor; the plot is negatively correlated with plural words in general on the x axis, and present tense on the y axis

Here's a dendrogram:

```{r}

rownames(gorman_clust) <- gorman.vars$title

# fviz_dend(hclust(dist(gorman_clust)),
#   palette = palette,
#   geom = geom,
#   ggtheme = ggtheme,
#   ellipse.type = ellipse.type,
#   cex = 0.6,
#   lwd = 0.1
# ) + dend_theme

hclust_gorman <- hclust(dist(gorman_clust), method = 'complete')
ggdendrogram(hclust_gorman)

ggsave(filename = "Images/dendrogram_gorman.png", dpi = "print")
```

```{r}


distance <- get_dist(df_scaled)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) + distance_labels + theme(axis.text = element_text(size = 2))

fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) + distance_labels + theme(axis.text = element_text(size = 2))

ggsave(filename = "Images/Distance plot with main d.f.png", dpi = "print")
```

## Gorman-style variables section-by-section

```{r K-means clustering-combined-sectioned}
set.seed(126126)

#Uncomment this when working with section numbers
gorman_clust <- sectioned.gorman.vars #|> filter(str_detect(title, "(spanish |gallic _ 1)"))

labs_kmeans <- gorman_clust$title

#gorman_clust <- gorman_clust |> rowwise() |> mutate(commentary = strsplit(title, ))

gorman_clust <- gorman_clust |> dplyr::select(-title)

kmeans_basic_cluster <- kmeans(gorman_clust, centers = k, nstart = 25)
kmeans_basic_table <- data.frame(kmeans_basic_cluster$size, kmeans_basic_cluster$centers)
kmeans_basic_df <- data.frame(Clusters = kmeans_basic_cluster$cluster, gorman_clust)

kmeans_basic_df <- kmeans_basic_df |> mutate(title = labs_kmeans) |> relocate(title, .before = 1)
```

Let's do our first visualization. Four clusters seems like a reasonable amount, looking at the fviz_nbclust results with the wss method.

```{r}
# TODO ADD SUMMARIES OF THE ABOVE DATA HERE

# Principal component analysis

pca <- prcomp(gorman_clust)

summary(pca)

# Visualizations

fviz_nbclust(gorman_clust, kmeans, method = "wss", nstart = 50)
#fviz_nbclust(gorman_clust, kmeans, method = "gap", nstart = 50, nboot = 50)
#fviz_nbclust(gorman_clust, kmeans, method = "silhouette", nstart = 50)

dim1 <- as.character(summary(pca)$importance[2] * 100)
dim2 <- as.character(summary(pca)$importance[5] * 100)

cluster_main <- fviz_cluster(kmeans_basic_cluster,
  data = gorman_clust,
  palette = palette,
  geom = geom,
  ellipse.type = ellipse.type,
  ggtheme = ggtheme
) +
  geom_text_repel(
    aes(label = labs_kmeans),
    vjust = "inward",
    hjust = "inward",
    max.overlaps = 14,
    size = 2,
    seed = 12927
  ) +
  labs(
    x = paste(strwrap(paste("Dim1", dim1, ""), 70),
      collapse = "\n"
    ),
    y = paste(strwrap(paste("Dim2", dim2, ""), 50), collapse = "\n"),
    title = "(1) K-means clustering applied to PHI Latin Texts, k= "
  ) +
  theme(axis.title = element_text(size = 12))

cluster_main


ggsave(filename = "Images/cluster-diagram-gorman-style-by-section.png", dpi = "print")

str(kmeans_basic_cluster)

```

The rotations suggest that the explanation of the data is poor; the plot is negatively correlated with plural words in general on the x axis, and present tense on the y axis

Here's a dendrogram:

```{r}
rownames(gorman_clust) <- sectioned.gorman.vars$title
gorman_clust$title <- sectioned.gorman.vars$title

ggdendrogram(hclust(dist(gorman_clust |> filter(str_detect(title, "(spanish|gallic _ 1)"))))) + theme(text = element_text(size = 6))

ggsave(filename = "Images/dendrogram_gorman.png", dpi = "print")
```

## Fuzzy Clustering

## CLIQUE Clustering

## Questions to Investigate

Which variables have the strongest correlation with these texts?

Are variations within the Gallic War enough to suggest episodic publication?

What does the cluster diagram look like when we go section by section?

## Pairwise similarity

First, let's add a new variable to the sectioned corpus which contains only the name of the commentary. The goal is, for each commentary, to get a pairwise similarity score of every vector using Spearman's ρ. That way, we can get a p-value for how likely the correlation is to be due to chance for each text, or not. We can also see which texts have high internal inconsistencies.

```{r}

#Add a 'path' variable which is just the commentary name
# TODO: Remove this cell, and use the title field instead? Or include this in the following cells

sectioned.gorman.vars %<>% 
  rowwise() %>%
  mutate(path = (strsplit(title, "_") |> unlist())[1:2] |> paste0(collapse = "_")) %>%
  relocate(path, .before = 2)

```

Now we'll create a list of vectors, and each one is the set of correlations retrieved by comparing pairwise

```{r}
if (!'pspearman' %in% .packages()) library(pspearman)
if (!'Kendall' %in% .packages()) library(Kendall)
if (!'lsa' %in% .packages()) library(lsa)

retrieve_combination_similarity <- function(combo, section_vectors, type) {
  vectors <-section_vectors |> filter(title %in% combo) |> dplyr::select(-title, -path)
  #browser()
  vec_a <- vectors[1,] |> t() |> as.vector() 
  vec_b <- vectors[2,] |> t() |> as.vector() 
  
  #browser()
  #If using Spearman, use approximation AS89 when the number of variables is under 1,290
  if (type == "spearman") {
    amount_jitter <- 0.001
    vec_a %<>% jitter(amount = amount_jitter)
    vec_b %<>% jitter(amount = amount_jitter)
    
    rho_estimate <- spearman.test(vec_a, vec_b, alternative = "two.sided", approximation = "AS89")$estimate
    #Note: we'll be comparing this with dissimilarity scores, which are actually the absolute value of (rho - 1)
    return(abs(rho_estimate - 1.0))  
  } else if (type == "cosine" ) {
    return(cosine(vec_a, vec_b))  
  } else if (type == "euclidean") {
    return(dist(rbind(vec_a, vec_b)))
  }
  
}

#Parameters:
#  character book: Name of commentary book. Valid values: "gallic _ 1", "gallic _ 2", etc.
#  tibble section_vectors: NEEDSDOC
# double prop: The proportion of sections to sample from to save on computation; should be 1 in the final article
within.book.similarity <- function(section_vectors, rows, type_of_similarity = "spearman") {
  
  sub_section_vectors <- section_vectors[sample(1:nrow(section_vectors), rows),]
  #browser()
  
  log_info("Starting sub_section_vectors...")
  t1 <- proc.time()
  
  #Use the columns if this to select
  #combn(sectioned.gorman.vars |> filter(path == "african ") |> pull(title), 2)
  combos <- combn(sub_section_vectors$title, 2)
  
  similarities <- apply(X = combos, MARGIN = 2, FUN = retrieve_combination_similarity, section_vectors = sub_section_vectors, type = type_of_similarity)
  
  print("Finished one book in within.book.similarity")
  print(proc.time() - t1)
  return(similarities)
  
}

```

Using Kendall, BA took 239.47 seconds to execute.

Using Spearman with the AS89 approximation, BA took 208.86 seconds to execute. We'll stick with that for now.

Let's create a function that does this all for us and gets us a list of vectors.

```{r}

#Get a vector of all the titles2
titles <- unique(sectioned.gorman.vars$path)

#Get the spearman correlation coefficients, add them to a list, with the label being the title


scores.correlation.per.text <- tapply(sectioned.gorman.vars, INDEX = sectioned.gorman.vars$path, FUN = within.book.similarity, rows = 16, type_of_similarity = "spearman")

names(scores.correlation.per.text) <- titles

gorman.vars$title <- lookup_table[gorman.vars$title]
gorman.vars <- as.data.frame(gorman.vars)
rownames(gorman.vars) <- gorman.vars$title
gorman_dist <- get_dist(gorman.vars |> dplyr::select(-title), method = "spearman")
#gorman_dist <- proxy::dist(gorman.vars |> dplyr::select(-title), method = "cosine")
#gorman_dist <- get_dist(gorman.vars |> dplyr::select(-title), method = "euclidean")
attributes(gorman_dist)$Label <- as.factor(attributes(gorman_dist)$Label)
attributes(gorman_dist)$Labels <- as.factor(attributes(gorman_dist)$Labels)

mat <- as.matrix(gorman_dist)
d_long <- expand.grid(Var1 = as.factor(rownames(mat)), Var2 = as.factor(colnames(mat)))
d_long$value <- as.vector(mat)
d_long$value <- round(d_long$value, digits = 2)
d_long$x_axis <- rep(1:length(gorman.vars$title), times = length(gorman.vars$title))
d_long$y_axis <- rep(1:length(gorman.vars$title), each = length(gorman.vars$title))
d_long <- as.data.frame(d_long)

graph <- fviz_dist(gorman_dist, show_labels = TRUE) + scale_x_discrete(labels = attributes(gorman_dist)$Labels) + scale_y_discrete(labels = attributes(gorman_dist)$Labels)
graph <- graph + 
  geom_tile(
    data = d_long,
    aes(x = x_axis, y = y_axis, fill = value), 
    inherit.aes = FALSE
            ) +
  geom_text(
              data = d_long,
              aes(x = x_axis, y = y_axis, label = value), 
              color = "black", 
              size.unit = "pt",
              size = 5,
              inherit.aes = FALSE) 

#Using the 5th percentile here. The 95th percentile is beyond any differences between texts, suggesting that difference doesn't guarantee different authorship. However, below about 0.2 is highly suggestive of same authorship, as that's lower than the lower tail (0.05) of the data
graph + scale_fill_gradient2(low = "red", high = "blue", midpoint = quantile(unlist(scores.correlation.per.text), probs = 0.05))
ggsave("Images/dissimilarity_matrix_between_books_by_correlation.png", dpi = "print")

graph + scale_fill_gradient2(low = "red", high = "blue", midpoint = 0.3)
ggsave("Images/dissimilarity_matrix_between_books.png", dpi = "print")

#View(as.data.frame(as.matrix(gorman_dist)))

```

Dissimilarity matrix with just Cicero, B Alex, and B Gall 5-8

```{r}
#be lazy and copy code

gorman.vars.2 <- gorman.vars |> filter(title %notin% c("BG 1", "BG 2", "BG 3", "BG 4", "BAlex. 1-21", "BHisp.", "BC 1", "BC 2", "BC 3", "BAfr.")) 
gorman_dist <- get_dist(gorman.vars.2 |> dplyr::select(-title), method = "spearman")
#gorman_dist <- proxy::dist(gorman.vars |> dplyr::select(-title), method = "cosine")
#gorman_dist <- get_dist(gorman.vars |> dplyr::select(-title), method = "euclidean")
attributes(gorman_dist)$Label <- as.factor(attributes(gorman_dist)$Label)
attributes(gorman_dist)$Labels <- as.factor(attributes(gorman_dist)$Labels)

mat <- as.matrix(gorman_dist)
d_long <- expand.grid(Var1 = as.factor(rownames(mat)), Var2 = as.factor(colnames(mat)))
d_long$value <- as.vector(mat)
d_long$value <- round(d_long$value, digits = 2)
d_long$x_axis <- rep(1:length(gorman.vars.2 $title), times = length(gorman.vars.2 $title))
d_long$y_axis <- rep(1:length(gorman.vars.2 $title), each = length(gorman.vars.2 $title))
d_long <- as.data.frame(d_long)

graph <- fviz_dist(gorman_dist, show_labels = TRUE) + scale_x_discrete(labels = attributes(gorman_dist)$Labels) + scale_y_discrete(labels = attributes(gorman_dist)$Labels)
graph <- graph + 
  geom_tile(
    data = d_long,
    aes(x = x_axis, y = y_axis, fill = value), 
    inherit.aes = FALSE
  ) +
  geom_text(
    data = d_long,
    aes(x = x_axis, y = y_axis, label = value), 
    color = "black", 
    size.unit = "pt",
    size = 5,
    inherit.aes = FALSE) 



graph + scale_fill_gradient2(low = "red", high = "blue", midpoint = quantile(unlist(scores.correlation.per.text), probs = 0.05), limits = c(0, 0.8))
ggsave("Images/dissimilarity_matrix_between_books_by_correlation-2.png", dpi = "print")

```

How about the mean dissimilarity and the Student's t-test that Li et al. used?

```{r}

#These indices currently only select Caesar
caesar_counts <- as_tibble(mat[17:30,17:30]) |> rowwise() |> mutate(means = mean(c_across(1:13)))
caesar_counts$title <- colnames(caesar_counts)[-15]

gallic_counts <- as_tibble(mat[22:29,22:29]) |> rowwise() |> mutate(means = mean(c_across(1:8)))
gallic_counts$title <- colnames(gallic_counts)[-9]
```

Perform a Student's t-test on each combination of texts in the dissimilarity matrix

```{r}

col1 <- rep(dimnames(mat)[[1]], times = length(dimnames(mat)[[1]]))
col2 <- rep(dimnames(mat)[[1]], each = length(dimnames(mat)[[1]]))

df <- tibble(col1 = col1, col2 = col2)

df %<>% rowwise() %>%
  mutate(
  pvalue = t.test(
    mat[unlist(col1),], 
    mat[unlist(col2),])$p.value
  )

df %<>% pivot_wider(names_from = col2, values_from = pvalue)
```
