---
title: "data_clustering"
output: html_document
date: "2026-04-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/PycharmProjects/corpus-caesarianum-authorship')
```


##Get the data

```{r}
final.frame <- read_csv("/feature_values.csv")
```


### Pre-visualization

I want to do a little bit of previsualization first.

Let's see what data type is in each column:

```{r}
as.data.frame(summary(final.frame))
```

Right, I forgot about the title column; let's find the mean of every
column except the title column.

```{r}
means <- final.frame |>
  tidytable::select(-title) |>
  apply(MARGIN=2, FUN=mean)

#get rid of scientific notation
options(scipen=999)

as.data.frame(sort(means))
```

Notice that a feature being common on a word isn't the same as it being
common on a parent. Not every part of speech is equally likely to be a
parent of another word. As a result, a parent being an auxiliary in a
passive verb is very uncommon, for example, coming close to the top of
the above list.

### Resume Data Prep

Now that we have a viable DF, let's cut it up a little bit. We don't
want the first column, because that's all text. Not every clustering
algorithm needs the DF to be pre-scaled (that is, subtracting the mean
from every value and dividing it by the standard deviation to make the
comparable). However, some do expect the data to be scaled, so lets
create two variables. **df_clust** isn't scaled, and **df_scaled** is.

```{r df_clust & df_scaled}
df_clust <- final.frame[, 2:length(final.frame)]
rownames(df_clust) <- final.frame$title
# df_clust <- slice_sample(df_clust, n=45)

df_scaled <- as.data.frame(scale(df_clust))
```

# Analysis

First, to get some of the formatting predefined.

Another note is that here is where I remove the variables I'm not
considering.

```{r}
# These vars I selected based on the Better variable selection heading down below.
# These are the variables that belonged to the 30 with the highest frequency AND the highest variance
# I also removed Act because it isn't independent from passive verbs, and I removed gender because there's no reason to believe it's related to register.
# vars <- c("Pres", "NOUN", "Imp", "2", "Past", "Perf", "3", "PRON", "Fin", "Nom", "Abl", "Prs", "Acc", "ADJ", "Gen", "ADP", "Part", "Pass", "DET", "Rel")
# 
# df_clust <- df_clust %>% dplyr::select(all_of(vars))
# rownames(df_clust) <- rownames(final.frame)
# df_scaled <- df_scaled %>% dplyr::select(all_of(vars))
# rownames(df_scaled) <- rownames(df_scaled)
```

```{r}


# Add this with the + operator to fviz_cluster plots. @param data is the original dataset used for cluster analysis (df_clust in the first case)
cluster_labels <- geom_text_repel(aes(label = rownames(data)), vjust = "inward", hjust = "inward", max.overlaps = 20, size = 3)
# Args for fviz_clust
palette <- c("#2E9FDF", "#00AFBB", "#E7B800", "red", "gray")
geom <- "point"
ellipse.type <- "convex"
ggtheme <- theme_bw()

# K-means
k <- 3

# Hcut
hc_func <- "agnes"
hc_method <- "ward"
hc_metric <- "euclidean"
cex <- 0.2

# Text labels for distance diagrams. @param label is usually the rownames of a data.frame
margin_size <- 10
distance_labels <- theme(axis.text = element_text(size = 5, vjust = 1), margins = margin(t = margin_size, r = margin_size, b = margin_size, l = margin_size))

# Dendrogram theme
dend_size <- 0.2
dend_theme <- theme(axis.text = element_text(size = dend_size, vjust = 1), margins = margin(t = margin_size, r = margin_size, b = margin_size, l = margin_size))
```

## K-means clustering

Time to cluster! K-means is a good starting point here. I'll check out
CLARA as well. We'll need to evaluate what's best. Here's the
modifications we can make:

1.  Change which variables we account for.

2.  Change the number of centers (between 3 and 5? But we already have a
    way of visualizing that)

3.  Change the measuring method (i.e. Euclidean

4.  Change the algorithm

```{r K-means clustering}
set.seed(126126)
kmeans_basic_cluster <- kmeans(df_clust, centers = k, nstart = 25)
kmeans_basic_table <- data.frame(kmeans_basic_cluster$size, kmeans_basic_cluster$centers)
kmeans_basic_df <- data.frame(Clusters = kmeans_basic_cluster$cluster, df_clust)
```

Let's do our first visualization. Four clusters seems like a reasonable
amount, looking at the fviz_nbclust results with the wss method.

```{r}
# TODO ADD SUMMARIES OF THE ABOVE DATA HERE

# Principal component analysis

pca <- prcomp(df_clust)

summary(pca)

# Visualizations

fviz_nbclust(df_clust, kmeans, method = "wss", nstart = 50)

dim1 <- as.character(summary(pca)$importance[2] * 100)
dim2 <- as.character(summary(pca)$importance[5] * 100)

cluster_main <- fviz_cluster(kmeans_basic_cluster,
  data = df_clust,
  palette = palette,
  geom = geom,
  ellipse.type = ellipse.type,
  ggtheme = ggtheme
) +
  geom_text_repel(
    aes(label = rownames(df_clust)),
    vjust = "inward",
    hjust = "inward",
    max.overlaps = 14,
    size = 2,
    seed = 12927
  ) +
  labs(
    x = paste(strwrap(paste("Dim1", dim1, "(Left = more 1st/2nd person, imperfective verbs, Right = more Ablative nouns, perfective verbs, and nouns overall)"), 70),
      collapse = "\n"
    ),
    y = paste(strwrap(paste("Dim2", dim2, "(Up = more 3rd person and past tense verbs, Down = more nouns)"), 50), collapse = "\n"),
    title = "(1) K-means clustering applied to PHI Latin Texts, k=4"
  ) +
  theme(axis.title = element_text(size = 12))

cluster_main


ggsave(filename = "Images/cluster-diagram-kmeans.png", dpi = "print")

str(kmeans_basic_cluster)

```

```{r}


distance <- get_dist(df_scaled)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) + distance_labels + theme(axis.text = element_text(size = 2))

fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07")) + distance_labels + theme(axis.text = element_text(size = 2))

ggsave(filename = "Images/Distance plot with main d.f.png", dpi = "print")
```

The graph above should be the first visualization in the presentation.
Next, we want to see what sets the Letters and other similar texts apart
from the others.

The following graph represents the loadings for the X axis of the
cluster analysis above. The farther we go to one side, the more 1st/2nd
person verbs, personal pronouns, present/future tense verbs, and
imperfective aspect verbs there are. Let's talk about what that means.

Person corresponds to a change in tenor accompanied by colloquial texts.
Tenor, as a term, refers to the author and their relationship to their
audience. Comedy, letters, and works of Seneca which come very close to
that cluster have an intimate audience (in the case of Comedy and
Seneca, this is an affectation. In such a close situation, the goal is
to both get information from the addressee but also communicate
information about themselves.

This information about person is fairly obvious, but the situation with
personal pronouns is less obvious. Latin has no need of personal
pronouns; the grammars say they tend to be used for emphasis or
...(*cite Allen & Greenough and Kühner / Stegmann*).

The information about tense is also expected. Speakers in colloquial
texts have a variety of purposes: asking and answering questions,
demonstrating their current thoughts and feelings, making plans, and so
on. These cover a broad range of statements, each of which requires a
different type of verb (*examples here*).

To sum up, texts on the left-hand side of the graph have a highly
personal tenor with ....FILL IN HERE, RE-READ BIBER

Texts on the the right-hand side of the cluster graph feature more
ablative nouns, participles, and passive verbs.

```{r}
pca <- prcomp(df_clust)

rotations.df <- as.data.frame(pca$rotation)

rotations <- data.frame(var = rownames(rotations.df), rotations = rotations.df$PC1)

rotation.two <- data.frame(var = rownames(rotations.df), rotations = rotations.df$PC2)

ggplot(rotations, aes(var, rotations)) +
  labs(x = "Variables", y = "Correlation Strength", title = "(2) Variation across X-axis in cluster diagram") +
  geom_col() +
  theme(
    axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1),
    axis.title = element_text(size = 14),
    title = element_text(size = 14)
  )
ggsave(filename = "Images/PC1_rotations.png", dpi = "print")

ggplot(rotations, aes(var, rotations)) +
  labs(x = "Variables", y = "Correlation Strength", title = "(2) Variation across X-axis in cluster diagram") +
  geom_col() +
  theme(
    axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1),
    axis.title = element_text(size = 14),
    title = element_text(size = 14)
  )
ggsave(filename = "Images/PC1_rotations.png", dpi = "print")

ggplot(rotation.two, aes(var, rotations)) +
  labs(x = "Variables", y = "Correlation Strength", title = "Correlation of Y-axis in cluster diagram with selected variables") +
  geom_col() +
  theme(
    axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1)
  ) +
  theme(
    axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1)
  )
ggsave(filename = "Images/PC2_rotations.png", dpi = "print")

# NOTE: I should try this with Act as a variable too. Pass alone might be high because there are a lot of participles and ablative nouns.
```

How different are the values of Atticus compared to the means of its own
cluster?

```{r}
# Lets get a vector of each of the Att values minus the mean of its cluster.

cluster_four <- kmeans_basic_df |> filter(Clusters == 1)

dist <- cluster_four |>
  dplyr::select(-title) |>
  map(
    \(x) x[1] - mean(x[-1])
  )

dist_df <- data.frame(var = names(dist)[-1], distance = unlist(dist)[-1])

ggplot(dist_df, aes(var, distance)) +
  geom_col() +
  scale_fill_continuous(breaks = c(0, 50, 100, 150, 200, 250)) +
  labs(title = "(3) Difference Between Letters to Atticus and Its Cluster") +
  xlab("Variable (X1 and X2 = 1st person and 2nd person)") +
  ylab("Letters to Atticus minus cluster mean") +
  theme(
    axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1),
    axis.title = element_text(size = 14),
    title = element_text(size = 14)
  )

ggsave(filename = "Images/att-vs-same-cluster-means.png", dpi = "print")
```

What about the differences between Comedy and the Letters to Atticus?

```{r}
kmeans_basic_df$title <- as.vector(final.frame$title)
cluster_four <- kmeans_basic_df |>
  filter(Clusters == 4 | title == "1 Letters to Att") |>
  dplyr::select(-title)

dist <- cluster_four |>
  map(
    \(x) x[1] - mean(x[-1])
  )

dist_df <- data.frame(var = names(dist)[-1], distance = unlist(dist)[-1])

ggplot(dist_df, aes(var, distance)) +
  geom_col() +
  scale_fill_continuous(breaks = c(0, 50, 100, 150, 200, 250)) +
  labs(title = "(4) Difference Between Letters to Atticus and Comedy") +
  xlab("Variable (X1 and X2 = 1st person and 2nd person)") +
  ylab("Letters to Atticus minus cluster mean") +
  theme(
    axis.text.x = element_text(size = 15, angle = 90, vjust = 0.5, hjust = 1),
    axis.title = element_text(size = 14),
    title = element_text(size = 14)
  )

ggsave(filename = "Images/atticus-vs-comedy-means.png", dpi = "print")
```

## Exploratory Visualization

We need to understand the variables before we go into much more detail.
Which ones are the most frequent? Which ones vary the most? Do we have a
lot of covariates? I'll establish a list of procedures in more detail
later, but, for now, just seeing the frequency of each variable by text
would be desirable

```{r}

```

## Gorman-Style Variables

```{r}
"pass"
```
