---
title: "corpus-caesarianum-authorship"
output:
  html_document:
    toc: true
    df_print: paged
editor_options:
  markdown:
    wrap: 72
params:
  use_att: TRUE
  use_parent: TRUE
  new_combos: TRUE
bibliography: rmd-notebook.bib
---

This notebook includes the code used to process the raw, postagged csv
data and extract the variables we'll be using to cluster the the texts
of the *Corpus Caesarianum*.

Several headings in this notebook are experiments or alternate methods
of variable selection not used in the final version of the article. They
are not labeled differently for now, as the project is still in
progress, but sections that are old or unnecessary are set to not
evaluate with the `knitr::` option `eval=FALSE`.

Some sections are currently unused, but I might revisit them. The
`use_att` param in the yaml header above lets us toggle the Simple
Variable Counts header on and off.

```{r include=TRUE}
knitr::opts_chunk$set(results = "hold")
#knitr::opts_knit$set(root.dir = '~/PycharmProjects/corpus-caesarianum-authorship')
#knitr::opts_knit$set(root.dir = '~/GitHub/corpus-caesarianum-authorship')
#knitr::opts_knit$set(root.dir = '/Users/matthewdehass/Documents/GitHub/corpus-caesarianum-authorship')
```

# Data Preparation

This section loads all the necessary libraries and prepares the `data`
variable for processing in the sections below. `data` draws from the
data in the ./postagged folder.

```{r, Libraries, results='hide'}

library(conflicted)
library(factoextra)
#library(dplyr)
library(ggrepel)
library(tidyverse)
library(cluster)
library(dendextend)
library(combinat)
library(MASS)
library(multcomp)
library(ggtext)
library(printr)
library(magrittr)
library(logrittr)
library(data.table)
library(collapse)
library(logger)

conflict_prefer("filter", "dplyr")
conflicted::conflicts_prefer(combinat::combn)
conflicted::conflicts_prefer(dplyr::select)

#For debugging purposes, here is a vector to save the size of the lower quartile that is discarded when variables are counted. This number is the minimum number of times per 1000 words a variable can occur in that section without getting discarded
lower_quartiles <- vector(mode = "double")

log_threshold(level = DEBUG)

source("functions.R")
```

Note that I was learning dplyr and tidyverse as I created this file, so
there are some confusions I should clear up here.

- tidyverse loads several other packages, including tibble

- tidytable uses tidyverse syntax over data.table

- dplyr is a syntax for working with data frames, including tibbles

- The `%>%` pipe is not identical to the `|>` pipe.

  - The former is from the magrittr package, the latter is native to R

  - The former lets you change the position of the piped-in argument by
    replacing an argument with a dot (.) R's `|>` allows this, but with
    \_ as the placeholder and the argument must be named

  - You can access members of a piped output with `%>%`, so
    `data.frame %>% mean(.$column)` works with magrittr but not base R.
    dplyr::pull() can do this with the base R pipe, though

(For the above ggrepel library, see
<https://github.com/slowkow/ggrepel>)

```{r Get postagged_only.csv. Big data!}

# dims <- c(
#   "title", 
#   "author", 
#   "book",
#   "section",
#   "path",
#   "form",
#   "lemma",
#   "tag",
#   "Aspect",
#   "Mood",
#   "Number",
#   "Person",
#   "Tense",
#   "VerbForm",
#   "Voice",
#   "Case",
#   "PronType",
#   "Gender",
#   "Polarity",
#   "Degree",
#   "NumType",
#   "Deprel",
#   "parent_form",
#   "parent_lemma",
#   "parent_tag",
#   "parent_Aspect",
#   "parent_Mood",
#   "parent_Number",
#   "parent_Person",
#   "parent_Tense",
#   "parent_VerbForm",
#   "parent_Voice",
#   "parent_Case",
#   "parent_PronType",
#   "parent_Gender",
#   "parent_Polarity",
#   "parent_Degree",
#   "parent_NumType",
#   "parent_Deprel"
# )

data <- fread("./postagged/postagged-texts.csv", na.strings = "", header = TRUE, stringsAsFactors = TRUE)

data <- bind_rows(
  data,
  fread("./postagged/postagged-cicero.csv", na.strings = "", col.names = colnames(data), stringsAsFactors = TRUE)[-1,],
  fread("./postagged/postagged-sallust.csv", na.strings = "", col.names = colnames(data), stringsAsFactors = TRUE)[-1,]
  )

data %<>%
    mutate(Person = as.factor(Person))

summary(data)
```

```{r processing-1, collapse=TRUE}
# Change some of the data types and rename certain values; details available in functions.R
data %<>% prepare_data()
summary(data)

```

### Removing Variables

The following section is based heavily in the analysis developed in the
[dashboard](./postagged-data-accuracy-dashboard.Rmd).

The first step is to combine features that are low accuracy or poorly
distinguished. The Rmd dashboard provided in this GitHub repo shows some
of the inaccuracy data.

## Simple Variable Counts

Now that we have the data, what we're *really* interested in is how many
of each tag appears in each work (how many of each POS, how many perfect
verbs, etc.). The point is, we're not interested in every individual
word.

NOTE: I've created a heading down below where we test formatting
variables like Gorman 2020. That is now the method we've adopted for our
most up-to-date results; this section can be safely skipped for the
moment.

This block of code will loop through and create a new data frame called
***final.frame*** (which is probably poorly named, but I'll fix it later
or I won't). This data frame has a variable for the count of each tag,
and every row is a unique work.

```{r Create final.frame}

#NOTE: There's a parameter for using these variables in the yaml header. By default, it is false and the next three cells (three if we include this one) will not run.

#NOTE 2: This currently only runs if we're using parent variables too
if (params$use_att && params$use_parent) {
  fdata <- data |> mutate(title = get_title_segment(title))
  titles <- unique(fdata$title)
  num_words <- table(fdata$title)
  final.frame <- data.frame(title = titles)
  
  for (i in 1:length(titles)) {
      frame <- fdata[fdata$title == titles[i],]
      frame <- tidytable::select(frame, title, tag:parent_Deprel & !parent_form & !parent_lemma) #Leave off everything unnecessary except for the title
      for (n in 2:length(frame)) {
        t <- table(frame[, ..n])
        # t <- prop.table(t)
        names <- names(t)
    
        for (name_index in seq(1, length(names))) {
          final.frame[i, names[name_index]] <- unname(t[names[name_index]])
        }
      }
  }
  
  # Normalize by number of words in each work
  for (i in 2:length(final.frame[1, ])) {
    for (n in 1:length(final.frame[, 1])) {
      final.frame[n, i] <- final.frame[n, i] * (10000 / num_words[[final.frame[n, ]$title]])
    }
  }
  
  # Removed punctuation because PCA showed that it was an "important" variable
  final.frame <- subset(final.frame, select = c(-PUNCT, -X))
  
  # Clean up NAs
  final.frame[is.na(final.frame)] <- 0
}
```

Let's save this to a file so we can open it in another RMarkdown file.

```{r save-simplex-vars}

if (params$use_att) write_csv(final.frame, file = "./feature_csv_files/feature_values.csv")
```

## Gorman Variables

Process is adapted from @gorman_author_2020

Step one of this process is to extract common feature pairs.

The number of feature values is `length(final.frame[1,]`.

That's a huge number, so let's start by working with the original data.

First, let's winnow accounting only for the main tags.

Let's get all combinations of two tags

```{r eval=FALSE}
library(combinat)

two_select_combinations <- combn(colnames(data |> tidytable::select(tag:parent_Deprel)), 2)

length(two_select_combinations)
```

Okay, that's 992 different combinations. We need to go through all the
combinations of variables, get the absolute frequency of each one. That
is not the frequency of each value (i.e., number of times the tag
`Masculine` occurs), but the number of times the feature itself occurs
(i.e., the number of words that have a value for the field `Gender`).
With the frequency information, we can keep only the high frequency
features as Gorman 2020 does. We use the same cutoff: the feature or
combination of features must appear in at least 20% of the corpus.

I've realized that counting up feature-value pairs at the end of this
file is proceeding too slowly, so I need to sample from the full dataset
for this process.

Note that the YAML header params `new_combos` and `use_parent` affect
the operation of the following code.

The data is reduced to only 1,000 tokens per book (so, 1,000 from BG
book 1, 1,000 from BG book 2, etc.; BAfr is one book, so only 1,000
tokens come from it).

The process is developed in hidden code blocks just below, which are
visible in the original RMarkdown file.

```{r, get-combinations-of-features, include=FALSE}

if (!exists("combinations")){
  combos_filename <- "combinations.csv"
  if (!params$new_combos && file.exists(combos_filename)) combinations <- read_csv(combos_filename) else {
  
  
  data_backup <- data

  #Weighting by book gives about a 50/50 split between Caesar and Cicero
  data <- data |>
    rowwise() |>
    #inelegant, but it works
    mutate(commentary = across(title, get_title_segment)) |>
    ungroup() |>
    dplyr::slice_sample(n = 1000, by = commentary)
  
  data$commentary <- as.factor(unlist(data$commentary))
  
  #Let's get every combination of variables.
  
  cols <- if (params$use_parent) colnames(data |> tidytable::select(tag:parent_Deprel & !parent_form & !parent_lemma)) else colnames(data |> tidytable::select(tag:Deprel))
  
  combinations <- bind_rows(
    tibble(a = cols), # Do it this way because we want to keep all of the single variables
    variable_combinations(2, data),
    variable_combinations(3, data),
    variable_combinations(4, data)
    #variable_combinations(5, data)
  )
  
  write_csv(combinations, file = combos_filename,)
  
  data <- data_backup
  
  rm(data_backup)
  }
}

```

If we need to re-run the data, we would go to this cell, run all
previous cells, and get started.

This cell uses the get_vars() function (which is in a hidden code block)
to take all the feature combinations retrieved above. For each
combination, it retrieves the frequency per 1,000 words of each value
for those features. So, the feature combination "Gender\|Number" would
return the normalized frequency of the following pairs of values:

|   Values   |
|:----------:|
| Masc, Sing |
| Fem, Sing  |
| Neut, Sing |
| Masc, Plur |
| Fem, Plur  |
| Neut, Plur |

```{r, entry-point}
#You can start the program here and hit "Run all chunks above" to 

if (!("future.apply" %in% (.packages()))) library(future.apply)
if (!("profvis" %in% (.packages()))) library(profvis)
if (!("dtplyr" %in% (.packages()))) library(dtplyr)

# profvis({

all_vars_by_section <- get_vars("chapter", data, presupplied_variables=NULL)
# })

# Ran a test of this function on the Bellum Hispaniense without culling 0-length variables.
# Result: 3,257,298 variables
# Test removing variables evidenced 0 times:
# Still 3,257,298. 
# Removing variables evidenced 1 or 0 times:
# 1,302,834
# Removing variables below the 25% quartile:
# Still 1,302,834, since the lower quartile is always 1
  
```

Each of these feature combinations is a column, each row a document or
document section, and the cells are the normalized frequencies within
those documents. However, the format the previous function returns is a
three-column data frame. One column is the variable combination, the
second is the document/document section title, and the third is the
frequency. We want to transpose this to match the arrangement just
described, which is more fitting for tidyverse functions and cluster
analysis. The following cells take care of this transposition.

```{r}
# Wait until we have all the results to we can select the top variables for ALL sections together
all_vars_by_section_culled <- all_vars_by_section |>
  select_top_variables(n = 1000) 
```

```{r}
all_vars_by_section_culled %<>% 
   transpose_all_vars()

if (exists("all_vars_by_section")) rm(all_vars_by_section)
gc() #free up memory for the OS

if (params$use_parent) filename <- "./feature_csv_files//feature_values_gorman_by_section.csv" else filename <- "./feature_csv_files//feature_values_wordonly_by_section.csv"

write_csv(all_vars_by_section_culled |> unnest(everything()), file = filename, col_names = TRUE)
```

The data above was for document sections, but we want to create
comparable data with full documents (books) so that we can compare. We
want to know, after all, if variation between documents is more than
that we see within documents, since we do not have many sure methods of
verifying our method's accuracy. The following function gets all the
variables and returns the set of feature combinations in the same format
we generated up above, so we can pass it to the get_vars() function
again.

```{r}
# Use the function from the above cell, get the variables, and then feed
# them back into the get_vars function to get the same variables, but with
# the frequencies calculated book-by-book this time.

combinations <- combinations_from_existing(all_vars_by_section_culled)
all_vars_by_book <- get_vars("book", data, combinations)

```

```{r}

# Transpose, and export to .csv file. N.B.: We don't cull uncommon
# variables, because we already are using a culled set of variables

all_vars_by_book <- all_vars_by_book %>%
  transpose_all_vars()

if (params$use_parent) filename <- "./feature_csv_files/feature_values_gorman.csv" else filename <- "./feature_csv_files/feature_values_wordonly.csv"

write_csv(all_vars_by_book, file = filename, col_names = TRUE)
```

## Data Verification

### Ensuring counts of variables are done correctly

This function takes a random cell from the final data frame and checks
its frequency in the main data frame.

```{r, eval=FALSE}
verify_random_cell

```

### Ensuring get_var_combos works correctly with different values

# Clustering

Clustering requires us to verify a number of conditions to ensure its
accuracy. Our targeted method is K-means (or successors like PAM), since
the clusters appear to be roughly spherical and our goal is to partition
the data. Given this focus, the following questions are pulled from
[@jain2010] p. 656:

> (a) What is a cluster?
>
> (b) What features should be used?
>
> (c) Should the data be normalized?
>
> (d) Does the data contain any outliers?
>
> (e) How do we define the pair-wise similarity?
>
> (f) How many clusters are present in the data?
>
> (g) Which clustering method should be used?
>
> (h) Does the data have any clustering tendency?
>
> (i) Are the discovered clusters and partition valid?

Multiple answers to each question must be tested, and not only to
determine the best method. We want to see that a variety of methods
generate similar partitioning, since that will boost the claim that the
texts belong to the categories they do. With variable selection
considered, we have made it to point (e), and we want a quick way to
consider many alternative of e-i. We will consider all alternatives
here, although the final article will only discuss the most salient or
list the rest in a footnote.

Our options for pair-wise similarity measures are the following:

- *Cosine distance (Deza and Deza 2009 Chs. 17, 19 and O Searcoid 2006)*

  $cosine(\theta) = \frac{v . w}{|v| |w|}$

- min-max measure [@kestemont_authenticating_2016]

- Cosine delta (Burrows 2002)

- Eder's Delta (Eder et al. 2016)

- SPATIUM-L1 [@kocher_unine_nodate] (Kocher 2016 at CLEF)

Determining the number of clusters can be done with the following
methods, each of which can get slightly different results, as there is
no one method. We must use our domain-specific knowledge to limit the
range:

- Gap statistic

- Total within sum of squares

- Average silhouette width

- SOM [@moisl2015]

- minimum message length (MML) [@jain2010] 656

- Gaussian mixture model (GMM) [@jain2010] 656

- minimum description length (MDL) [@jain2010] 656

- Bayes information criterion (BIC) [@jain2010] 656

- the Dirichlet process (DP) [@jain2010] 656

Which clustering method should be used is another non-trivial problem.
K-means has been expanded many different ways. Note [@jain2010] p. 659
cites Fisher and vanNess on the following:

1.  **Convex**: A clustering algorithm is convex-admissible if it
    results in a clustering where the convex hulls of clusters do not
    intersect.
2.  **Cluster proportion**: A clustering algorithm is cluster-proportion
    admissible if the cluster boundaries do not alter even if some of
    the clusters are duplicated an arbitrary number of times.
3.  **Cluster omission**: A clustering algorithm is omission-admissible
    if by removing one of the clusters from the data and re-running the
    algorithm, the clustering on the remaining K  1 clusters is
    identical to the one obtained on them with K clusters.
4.  **Monotone**: A clustering algorithm is monotone-admissible if the
    clustering results do not change when a monotone transformation is
    applied to the elements of the similarity matrix.

Here are a few options from Moisl:

- K-means

- Also PAM and CLARA\

- SOM (see Moisl notes below)\

- Agglomerative clustering\

- CLIQUE

- Dbscan (see Moisl pp. 166ff.)

- FANNY or Fuzzy clustering (see Kaufman and Rousseeuw 1990, chapter 4)

Does the data have clustering tendency is a question discussed in
[@moisl2015] and [@jain2010].
