---
title: "corpus-caesarianum-authorship"
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook
editor_options:
  markdown:
    wrap: 72
params:
  use_att: FALSE
bibliography: rmd-notebook.bib
---

This notebook includes the code used to process the raw, postagged csv data and extract the variables we'll be using to cluster the the texts of the *Corpus Caesarianum*.

Several headings in this notebook are experiments or alternate methods of variable selection not used in the final version of the article. They are not labeled differently for now, as the project is still in progress, but sections that are old or unnecessary are set to not evaluate with the `knitr::` option `eval=FALSE`.

Some sections are currently unused, but I might revisit them. The `use_att` param in the yaml header above lets us toggle the Simple Variable Counts header on and off.

```{r include=TRUE}
knitr::opts_chunk$set(results = "hold")
#knitr::opts_knit$set(root.dir = '~/PycharmProjects/corpus-caesarianum-authorship')
#knitr::opts_knit$set(root.dir = '~/GitHub/corpus-caesarianum-authorship')
#knitr::opts_knit$set(root.dir = '/Users/matthewdehass/Documents/GitHub/corpus-caesarianum-authorship')
```

# Data Preparation

This section loads all the necessary libraries and prepares the `data` variable for processing in the sections below. `data` draws from the data in the ./postagged folder.

```{r, Libraries, results='hide'}

library(conflicted)
library(factoextra)
library(dplyr)
library(ggrepel)
library(tidyverse)
library(cluster)
library(combinat)
library(dendextend)
library(MASS)
library(multcomp)
library(ggtext)
library(printr)
library(magrittr)
library(logrittr)
library(data.table)
library(collapse)
library(logger)

conflict_prefer("filter", "dplyr")
conflicted::conflicts_prefer(combinat::combn)
conflicted::conflicts_prefer(dplyr::select)

#For debugging purposes, here is a vector to save the size of the lower quartile that is discarded when variables are counted. This number is the minimum number of times per 1000 words a variable can occur in that section without getting discarded
lower_quartiles <- vector(mode = "double")

# 
# library(duckplyr)
# conflict_prefer("filter", "dplyr")
# Sys.setenv(DUCKPLYR_FALLBACK_INFO = TRUE)
# 
# conflicts_prefer(dplyr::filter)
```

Note that I was learning dplyr and tidyverse as I created this file, so there are some confusions I should clear up here.

- tidyverse loads several other packages, including tibble

- tidytable uses tidyverse syntax over data.table

- dplyr is a syntax for working with data frames, including tibbles

- The `%>%` pipe is not identical to the `|>` pipe.

  - The former is from the magrittr package, the latter is native to R

  - The former lets you change the position of the piped-in argument by replacing an argument with a dot (.) R's `|>` allows this, but with \_ as the placeholder and the argument must be named

  - You can access members of a piped output with `%>%`, so `data.frame %>% mean(.$column)` works with magrittr but not base R. dplyr::pull() can do this with the base R pipe, though

(For the above ggrepel library, see <https://github.com/slowkow/ggrepel>)

```{r Get postagged_only.csv. Big data!}

dims <- c(
  "title", 
  "author", 
  "book",
  "section",
  "path",
  "form",
  "lemma",
  "tag",
  "Aspect",
  "Mood",
  "Number",
  "Person",
  "Tense",
  "VerbForm",
  "Voice",
  "Case",
  "PronType",
  "Gender",
  "Polarity",
  "Degree",
  "NumType",
  "Deprel",
  "parent_form",
  "parent_lemma",
  "parent_tag",
  "parent_Aspect",
  "parent_Mood",
  "parent_Number",
  "parent_Person",
  "parent_Tense",
  "parent_VerbForm",
  "parent_Voice",
  "parent_Case",
  "parent_PronType",
  "parent_Gender",
  "parent_Polarity",
  "parent_Degree",
  "parent_NumType",
  "parent_Deprel"
)

data <- bind_rows(
  fread("./postagged/postagged-texts.csv", na.strings = "", col.names = dims, stringsAsFactors = TRUE),
  fread("./postagged/postagged-cicero.csv", na.strings = "", col.names = dims, stringsAsFactors = TRUE)
)

#This isn't a ratio or count variable
data %<>%
  mutate(Person = as.factor(Person))
# cicero_data %<>%
#   mutate(Person = as.factor(Person))

summary(data)
```

```{r}
#The Bellum Alexandrinum may represent two clear stages of composition, and we might want to respond to Gaertner's argument that the completion of Alex. and the BC. were one project. I'll do this by making sections 1-21 "book" 1 and the rest "book" 2

#Don't forget the sections are zero-indexed, so 0-20 is the supposedly Caesarian section, not 1-21.
#since I'm using a data table, subset without the comma, unlike base R
data$book[data$path == "alexandrine" & data$section > 20] <- 2
```

The data needs to be split up by section first, then put back into books. The way I recommend doing this is appending both the book and section number to the title to get this result: "*title*\_*x_x(x)*". Then, we use the same code blocks below as in the Letters to Atticus, feeding this new title instead of the base one.

Let's modify the titles:

```{r}

data <- data |>
  rowwise() |>
  mutate(title = gsub(" ", "", paste(title, "_", book, "_", section)))

# cicero_data <- cicero_data |>
#   rowwise() |>
#   mutate(title = paste(title, "_", book, "_", section))
```

The following two code blocks come from the [cltk-2025-atticus repository](https://github.com/cubis182/cltk-2025-atticus "GitHub Repo"), and they transpose the data so that each column is a feature value instead of a feature type and the values in the columns are the number of times each feature appears.

```{r processing-1}

# # Rename "Imp" to "Imperative" in mood category
# levels(data$Mood) <- c(levels(data$Mood), "Imperative")
# sequence <- which(data$Mood == "Imp")
# data$Mood[sequence] <- "Imperative"
# remove(sequence)
# 
# data$Mood[data$Mood == "Imp"] <- "Imperative"
# data$Mood <- droplevels(data$Mood) # Remove the old "Imp

#Turn it into a function
replace_value <- function(column, to_replace, replace_string) {
  # Rename "Imp" to "Imperative" in mood category
  levels(data[[column]]) <- c(levels(data[[column]]), replace_string)
  sequence <- which(data[[column]] == to_replace)
  data[[column]][sequence] <- replace_string
  remove(sequence)
  
  data[[column]][data[[column]] == to_replace] <- replace_string
  data[[column]] <- droplevels(data[[column]]) # Remove the old "Imp
  return(data)
}

#Caesar
data <- replace_value("Mood", "Imp", "Imperative")
data <- replace_value("Mood", "Ind", "Indicative")
data <- replace_value("parent_Mood", "parent_Ind", "parent_Indicative")
data <- replace_value("parent_Mood", "parent_Imp", "parent_Imperative")

#Cicero
# cicero_data <- replace_value("Mood", "Imp", "Imperative")
# cicero_data <- replace_value("Mood", "Ind", "Indicative")
# cicero_data <- replace_value("parent_Mood", "parent_Ind", "parent_Indicative")
# cicero_data <- replace_value("parent_Mood", "parent_Imp", "parent_Imperative")

#Remove unneeded variables
data %<>% tidytable::select(-NumType, -parent_NumType, -Degree, -parent_Degree) |>
  filter(tag != 'PUNCT')

# cicero_data %<>% tidytable::select(-NumType, -parent_NumType, -Degree, -parent_Degree) |>
#   filter(tag != 'PUNCT')

#An argument could be made for removing Polarity, but I'll wait to see how common it is and how much variance there is between the texts

```

## Utilities

Getting the book number from the string format *title_booknum_sectionnum*

```{r}

#modes: "title", "book", "section"
get_title_segment <- function(string, mode = "book") {
  split_str <- unlist(strsplit(string, "_"))

  #We use gsub to ensure there are no spaces, might be unnecessary
  gsub(" ", "", paste(
    split_str[1], 
    if (mode != "title") paste("_", split_str[2]), 
    if (mode == "section") paste("_", split_str[3])
    )
    )
}
```

## Simple Variable Counts

Now that we have the data, what we're *really* interested in is how many of each tag appears in each work (how many of each POS, how many perfect verbs, etc.). The point is, we're not interested in every individual word.

NOTE: I've created a heading down below where we test formatting variables like Gorman 2020. That is now the method we've adopted for our most up-to-date results; this section can be safely skipped for the moment.

This block of code will loop through and create a new data frame called ***final.frame*** (which is probably poorly named, but I'll fix it later or I won't). This data frame has a variable for the count of each tag, and every row is a unique work.

```{r Create final.frame}

#NOTE: There's a parameter for using these variables in the yaml header. By default, it is false and the next three cells (three if we include this one) will not run.

if (params$use_att) {
  titles <- unique(data$title)
  num_words <- table(data$title)
  final.frame <- data.frame(title = titles)
  
  for (i in 1:length(titles)) {
      frame <- data[data$title == titles[i],]
      frame <- tidytable::select(frame, title, tag:parent_Deprel & !parent_form & !parent_lemma) #Leave off everything unnecessary except for the title
      for (n in 2:length(frame)) {
        t <- table(frame[, ..n])
        # t <- prop.table(t)
        names <- names(t)
    
        for (name_index in seq(1, length(names))) {
          final.frame[i, names[name_index]] <- unname(t[names[name_index]])
        }
      }
  }
  
  # Normalize by number of words in each work
  for (i in 2:length(final.frame[1, ])) {
    for (n in 1:length(final.frame[, 1])) {
      final.frame[n, i] <- final.frame[n, i] * (10000 / num_words[[final.frame[n, ]$title]])
    }
  }
  
  # Removed punctuation because PCA showed that it was an "important" variable
  final.frame <- subset(final.frame, select = c(-PUNCT, -X))
  
  # Clean up NAs
  final.frame[is.na(final.frame)] <- 0
}
```

Let's save this to a file so we can open it in another RMarkdown file.

```{r save-simplex-vars}

if (params$use_att) write_csv(final.frame, file = "./feature_csv_files/feature_values.csv")
```

## Gorman Variables

Process is adapted from @gorman_author_2020

Step one of this process is to extract common feature pairs.

The number of feature values is `length(final.frame[1,]`.

That's a huge number, so let's start by working with the original data.

First, let's winnow accounting only for the main tags.

Let's get all combinations of two tags

```{r eval=FALSE}
library(combinat)

two_select_combinations <- combn(colnames(data |> tidytable::select(tag:parent_Deprel)), 2)

length(two_select_combinations)
```

Okay, that's 992 different combinations. We need to:

1.  Turn the pairwise_combination matrix into a tibble

    ```         
    two_select_combinations <- as_tibble(two_select_combinations)
    ```

2.  Apply the following to every combination in the list

3.  Count the number of rows in the data frame that don't have **NA** in either category

    - Note that selecting those two rows is as easy as the following line of code:

      ```         
      data |> tidytable::select(pairwise_combinations[,1])
      ```

    - Just plug in the column number and the output works in the select function

4.  In that column of the pairwise_combinations tibble, add a third field with the count retrieved in the previous step.

5.  Sort the tibble by the third field, cull out a percentage based on Gorman 2016Okay, so, we've gotten data on all the combinations of two variables, but what about 3 to 5? Let's refactor the above code into a function we can re-use:

```{r}

if (!("combinat" %in% (.packages()))) {
  library(combinat)
}

#Filters to rows which have ZERO NA values in them
no_na <- function(tib, vec) {
  if(NA %in% vec) {
    return(tib)
  }
  tib <- filter(tib, .data[[vec[1]]] != FALSE)
  vec <- vec[2:length(vec)]
  return(no_na(tib, vec))

}

#Given a matrix which includes a set of column names in the source_data, return the number of rows that have valid data in all columns
variable_count <- function(matrix, source_data) {
  
  subset_of_data <- source_data %>% 
    tidytable::select(all_of(matrix)) %>%
    no_na(colnames(.))
  
  return(nrow(subset_of_data)) 
    
}

#combinations_n <- 2

#Get rid of values that don't appear in more than 20% of the corpus
cull <- function(combinations) {
  print(paste("Length before culling: ", as.character(length(combinations$a))))
  
  #Get rid of features which occur in less than 20% of the corpus
  filtered <- combinations |> filter(num > (nrow(data) / 5))
  
  print(paste("Length after culling: ", as.character(length(filtered$a))))
  return(filtered)
}

#This function returns a list of all combinations of a variable which occur 
variable_combinations <- function(combinations_n, source_data) {
  select_combinations <- combn(colnames(source_data |> tidytable::select(tag:parent_Deprel & !parent_form & !parent_lemma)), combinations_n)
  
  #length(select_combinations)
  
  #Get a table showing the number of times all the features in a combination are attested
  combination_freqs <- apply(X=select_combinations, MARGIN = 2, FUN=variable_count, source_data = source_data)
  
  tib_select_combinations <- as_tibble(select_combinations)
  
  #Get the length of the column
  col_len <- nrow(tib_select_combinations)
  
  #Add a new row with the frequency each one occurs
  tib_select_combinations[(col_len + 1),] <- as.list(as.character(combination_freqs))
  
  #Time to take out the trash!
  remove(col_len)
  
  #transpose
  #browser()
  
  tib_select_combinations <-
    tib_select_combinations |> 
    #The number of columns depends on the $ of combinations, hence the code in the following line
    mutate(type = c(letters[1:combinations_n], 'num')) |> 
    #Don't select 'num'
    pivot_longer(!type) |> 
    pivot_wider(names_from = type, values_from = value) |> 
    dplyr::select(!name) |>
    mutate(across(num, as.integer))
  
  #browser()
  
  arrange(tib_select_combinations, num)
  
  tib_select_combinations <- tib_select_combinations |> filter(num > 0) |> arrange(num) 
  
  return(tib_select_combinations |> cull() |> dplyr::select(-num))
}


```

I've realized that counting up feature-value pairs at the end of this file is proceeding too slowly, so I need to sample from the full dataset.

```{r}

if (!exists("combinations")){
  combos_filename <- "combinations.csv"
  if (file.exists(combos_filename)) combinations <- read_csv(combos_filename) else {
  
  
  data_backup <- data

  #Weighting by book gives about a 50/50 split between Caesar and Cicero
  data <- data |>
    rowwise() |>
    #inelegant, but it works
    mutate(commentary = across(title, get_title_segment)) |>
    ungroup() |>
    dplyr::slice_sample(n = 1000, by = commentary)
  
  data$commentary <- as.factor(unlist(data$commentary))
  
  #Let's get every combination of variables.
  
  combinations <- bind_rows(
    variable_combinations(1, data),
    variable_combinations(2, data),
    variable_combinations(3, data),
    variable_combinations(4, data),
    variable_combinations(5, data)
  )
  
  write_csv(combinations, file = combos_filename,)
  
  data <- data_backup
  
  rm(data_backup)
  }
}

```

Let's create a function that takes each combination in the list and gets the number of times each one occurs. This function also removes any that occur zero times.

```{r}

#Requires a row from a tibble that only includes valid column names and no numerics / other data types
#Discards any data points that occur less than 30 times

get_var_combos <- function(var_combn, source_data) {
  #browser()
  
  #This is used to select the desired columns
  selection_char <- as.character(var_combn)
  selection_char <- selection_char[!is.na(selection_char)]
  
  var_combos <- source_data[selection_char]
  #browser()
  #Added this line to avoid the expensive na.omit() operation
  var_combos <-  var_combos[complete.cases(var_combos),]
  
  var_combos <- collapse::fcount(var_combos, drop = TRUE) #|> na.omit()
    #Added next line to avoid expensive string comparisons
  
  if (nrow(var_combos) > 0) {
    #final.table <- table(var_combos) #|> as_tibble()
    
    #browser()
    
    return(var_combos) #|> filter(n > 30)) #initially not 30, but nrow(source_data / 5)
    }
  else {return(NULL)}
}
```

This function takes a set of variable combinations (one_combinations, two_combinations, etc.) and applies the get_var_combos function to all of them, creating a data frame out of it.

```{r}


count_var_combos <- function(var_combn, source_data) {
  
  list_combn <- apply(var_combn, MARGIN = 1, get_var_combos, source_data = source_data) |> bind_rows()
  
    #unlist()
  
  #browser()
  #list_combn <- list_combn
  
  # browser()
  #Initially removed low-frequency items, but that resulted in NA values in the final data.
  #I've decided, though, that the reduction in quality is worth the increase in performance and lower memory usage. I remove anything below the lower quartile
  lower_quartile <- quantile(list_combn$N)[['25%']]
  .GlobalEnv$lower_quartiles <- c(.GlobalEnv$lower_quartiles, lower_quartile)
  return(list_combn[list_combn$N > lower_quartile,]) #subset(list_combn, list_combn > nrow(source_data) / 10)) # |> as_tibble() |> filter(V1 > 30))
}

#count_var_combos(two_combinations, data)
```

Some helper functions for the upcoming run_combos and get_vars function which creates the final data frame are in the cell below. They're a refactoring of the above cell, which should get replaced soon.

```{r}
convert_row_to_character <- function(row_of_factors) {
  row_of_factors |> mutate(across(where(is.factor), as.character)) 
}

has_all_feature_types <- function(row_from_source_data, character_vector_of_feature_values) {
  
  # Unnecessary when using apply() around this function
  # row_from_source_data %<>% mutate(across(everything(), as.character))
  
  # if (!is.character(row_from_source_data)) {
  #   row_from_source_data %<>% mutate(across(where(is.factor), as.character)) 
  # } 
  
  # Only return true for this row if ALL characters in the vector on the left are in the row on the right
  # Note on optimization: %chin% is supposedly faster than %in%
  all(character_vector_of_feature_values %in% row_from_source_data)
}

get_number_rows_with_feature_types <- function(vector_of_feature_values, source_data) {
  #browser()
  fn_name <- "get_number_rows..."

  count_of_feature_values <- sum(
    apply(X = source_data, MARGIN = 1, FUN = has_all_feature_types, character_vector_of_feature_values = vector_of_feature_values)
  )
  log_debug("{fn_name}: The feature values '{paste(vector_of_feature_values)}' occurred {count_of_feature_values} times in the dataset.")
  
  count_of_feature_values
}

```

Okay, now we create functions used down below for transposing the data (ensuring rows are observations and columns are variables) and selecting the top *n* variables by relative frequency.

```{r}

#all_vars <- all_vars |> mutate(title = unlist(title)

#First, like Gorman, remove ones appearing in less than 20% of the corpus. For section-by-section, that's 134 / 5 = 27

#Recommended operation on the output of get_vars()
transpose_all_vars <- function(all_vars) {
  all_vars <- all_vars |> mutate(title = unlist(title))
  
  all_vars <- pivot_wider(
    all_vars, 
    names_from = pasted, 
    values_from = n,
    values_fill = 0)
  
  return(all_vars |> as_tibble())
}

#Reduces output of transpose_all_vars to top n variables; maximum value for n is 18,000
select_top_variables <- function(all_vars, n) {
  #For logging
  fn_name <- get_logger_meta_variables(log_level = INFO)$fn
  
  #An optimization step to reduce the size of the data we're dealing with
  number_of_variables <- length(unique(all_vars$pasted))
  number_of_rows <- nrow(all_vars)
  min_val <- sort(all_vars$n, decreasing = TRUE)[1:(n * (number_of_rows / number_of_variables))] |>
    min()
  
  log_info("{fn_name}: Number of variables is {number_of_variables}, minimum value in sections 1:{(n * (number_of_rows / number_of_variables))} is {min_val}.")
  
  original_length <- nrow(all_vars)
  all_vars_reduced <- all_vars %>=% filter(n > min_val)
  log_info("{fn_name}: Data frame reduced from {original_length} rows to {nrow(all_vars_reduced)}.")
  
   # Get a list of the distinct variables
  variables <- unique(all_vars_reduced$pasted)
  log_info("{fn_name}: Number of variables to compare is {format(length(variables), big.mark = ',')}")
  
  # NOTE: COMMENTING THIS OUT FOR NOW, TO SEE HOW THE HEURISTIC PERFORMS
  # For each distinct variable, sum up its values across all works, then store in a list
  combined_vars <- lapply(variables, FUN = function(var) sum(all_vars_reduced$n[all_vars_reduced$pasted == var]))
  combined_vars %<>% unlist

  #Add the names to the list
  names(combined_vars) <- variables

  sorted <- sort(combined_vars, decreasing = TRUE)[1:n]
  to_filter_by <- names(sorted)
  
  all_vars |> filter(pasted %in% to_filter_by)
}

# title_temp <- all_vars$title
# all_vars <- all_vars |> dplyr::select(-title) |> dplyr::select(where(~ sum(.x, na.rm = TRUE) > length(.x) * 30)) |> mutate(title = title_temp)
# 
# all_vars <- all_vars |> relocate(title, .before = everything())
# 
# remove(title_temp)
```

Now, lets create a function that gets one long tibble with every single variable in it.

```{r}
#library(dtplyr)

#
#character select : A character value which is the section/book we're using in this function
count_combinations_per_section <- function(combinations, source_data, title_str) {
  #browser()
  t1 <- proc.time()
  norm_val <- 1000 / nrow(source_data)
  
  #save processing time by converting source_data to matrix now
  source_data <- as.matrix(source_data)
  
  #more efficient than using rows
  combinations <- t(as.matrix(combinations))
  
  counts <- apply(combinations, MARGIN = 2, FUN = get_number_rows_with_feature_types, source_data = source_data)
  counts <- counts * norm_val
  return_value <- bind_cols(t(combinations), tibble(n = counts))
  return_value$title <- title_str
    # dplyr::rowwise() %>%
    # summarize(
    #   n = get_number_rows_with_feature_types(c_across(everything()), source_data)
    # ) %>%
    # bind_cols(combinations, .) |>
    # mutate(n = n * norm_val, title = title_str)
  
  #browser()
  print(proc.time() - t1)
  return_value
}

run_combos <- function(source_data, title_str) {
  t1 <- proc.time()
  
  all_vars <- count_var_combos(combinations, source_data) |>
    unite(col = "pasted", -N, sep = "|", na.rm = TRUE) |>
    as_tibble(.name_repair = "universal_quiet")  
  
  #browser()

  names(all_vars)[1] <- "pasted"
  
  #browser()
  
  #Normalize frequencies by occurrence per 1000 words
  norm_val <- 1000 / nrow(source_data)
  all_vars <- all_vars |> 
    mutate(n = N * norm_val) |>
    dplyr::select(-N)
  
  all_vars <- all_vars |>
    mutate(title = title_str)  #The titles should all be the same...

  
  delta_time <- proc.time() - t1
  log_debug(paste(names(delta_time), delta_time, sep = ": ", collapse = ", "))
  return(all_vars)
}

#mode: whether to divide the data into commentaries, books, or chapters
#If getting the variables from an existing data frame, make sure they're
# in the combinations variable and use_preset is set to TRUE
get_vars <- function(mode = "book", source, presupplied_variables = NULL) {
  
  if (mode == "book") {
    #browser()
    source_data <- source |>
      rowwise() |>
      mutate(title = get_title_segment(title))
  }
  else if (mode == "commentary") {
    source_data <- source |>
      rowwise() |>
      mutate(title = get_title_segment(title, "title"))
  }
  else if (mode == "chapter") {
    source_data <- source 
  }
  else {
    return(NULL)
  }
  
  if (is.null(presupplied_variables)) {
    all_vars <- source_data |>
        group_by(title) %>%
        #lazy_dt() |>
        group_map(~ run_combos(.x, .y)) |>
        bind_rows() |>
        mutate(title = unlist(title))
  } else {
    #browser()
    all_vars <- source_data |>
      #For debugging:
      #slice_sample(n = 100) |>
      rowwise() |>
      #check whether this line is necessary
      #mutate(title = unlist(strsplit(title, "_"))[1:2] |> paste(collapse="_")) |>
      group_by(title) %>%
      group_map(~ count_combinations_per_section(presupplied_variables, source_data = .x, title_str = .y)) |>
      bind_rows() |>
      mutate(title = unlist(title)) |>
      unite(col = "pasted", 1:5, sep = "|", na.rm = TRUE)
  }
  
  return(all_vars)
}
```

If we need to re-run the data, we would go to this cell, run all previous cells, and get started.

```{r, entry-point}
#You can start the program here and hit "Run all chunks above" to 

if (!("future.apply" %in% (.packages()))) library(future.apply)
if (!("profvis" %in% (.packages()))) library(profvis)
if (!("dtplyr" %in% (.packages()))) library(dtplyr)

# profvis({

all_vars_by_section <- get_vars("chapter", data, presupplied_variables=NULL)
# })

# Ran a test of this function on the Bellum Hispaniense without culling 0-length variables.
# Result: 3,257,298 variables
# Test removing variables evidenced 0 times:
# Still 3,257,298. 
# Removing variables evidenced 1 or 0 times:
# 1,302,834
# Removing variables below the 25% quartile:
# Still 1,302,834, since the lower quartile is always 1
  
```

```{r}
# Wait until we have all the results to we can select the top variables for ALL sections together
all_vars_by_section_culled <- all_vars_by_section |>
  select_top_variables(n = 1000) 
```

```{r}
all_vars_by_section_culled %<>% 
   transpose_all_vars()

if (exists("all_vars_by_section")) rm(all_vars_by_section)
gc() #free up memory for the OS
write_csv(all_vars_by_section_culled |> unnest(everything()), file = "./feature_csv_files//feature_values_gorman_by_section.csv", col_names = TRUE)
```

We need to create a function that takes a set of columns and returns the same value that `variable_combinations()` does.

NOTE: This could be refactored into the get_vars() function, and probably should be. It also needs to be documented in more detail

```{r}

#Pass an output from get_vars(), get the variables back in table form so they can be passed into the get_vars() function with a new dataset. That way, a new dataset can be processed with the same variables as another.
combinations_from_existing <- function(all_vars) {
  #Get the names() attribute
  vars_unsplit <- all_vars |> dplyr::select(-title) |> colnames()
  
  
  #lapply over the names, using strsplit on each one
  split_vars <- lapply(X = vars_unsplit, FUN = strsplit, split="[|]")
  
  split_vars <- unlist(split_vars, recursive = FALSE)
  
  #For each item in the list (which is a character vector), we construct a data frame where each column is the character vector. NAs should appear in rows that aren't attested
  combinations_new <- data.frame(labels = letters[1:5])
  
  for (item in split_vars) {
    combinations_new <- cbind(combinations_new, item[1:5])
  }
  
  #Transpose the data frame
  combinations_new <- t(combinations_new) |> as_tibble(.name_repair = 'unique') |> slice(-1)
  
  combinations_new
}

```

Use the function from the above cell, get the variables, and then feed them back into the get_vars function to get the same variables, but with the frequencies calculated book-by-book this time.

```{r}
combinations <- combinations_from_existing(all_vars_by_section_culled)
all_vars_by_book <- get_vars("book", data, combinations)


#post-processing
# cic_all_vars_top <- cic_all_vars
#   transpose_all_vars() |>
#   select_top_variables(1000)
```

Transpose, and export to .csv file. N.B.: We don't cull uncommon variables, because we already are using a culled set of variables

```{r}

all_vars_by_book <- all_vars_by_book %>%
  transpose_all_vars()

write_csv(all_vars_by_book, file = "./feature_csv_files/feature_values_gorman.csv", col_names = TRUE)
```

We need to be able to reselect small segments of the corpus and count the same variables, so we need a function to find and count each column again.

## Data Verification

### Ensuring counts of variables are done correctly

This function takes a random cell from the final data frame and checks its frequency in the main data frame.

```{r}
if ("stringr" %notin% .packages()) {
  library(stringr)
}

verify_random_cell <- function(all_vars, source_data) {
  # For logging purposes
  fn_name <- get_logger_meta_variables(log_level = INFO)$fn
  
  # Select a random cell from all_vars
  rand_row <- slice_sample(all_vars)
  column_number <- sample(2:ncol(rand_row), size = 1)# column number of the targeted cell, excluding titles
  cell <- rand_row[,column_number]
  
  # Get its title and column name
  title <- rand_row$title
  colname <- names(all_vars)[ column_number ]
  
  log_info("{fn_name}: Selecting all_vars[title == {title}, '{colname}']")
  log_info("{fn_name}: Value of cell is {cell}")
  
  # Remove space from the title
  title_selected <- gsub(pattern = " ", replacement = "", title)
  
  # Split column name into a character vector
  vars_unsplit <- unlist(strsplit(colname, "[|]"))
  
  # Filter source_data to the title, using partial matches with str_starts() from stringr package
  filtered_data <- source_data |> filter(str_starts(gsub(" ", "", title), title_selected))
  
  # Pass character vector and source_data to get_number_rows_with_feature_types()
  count_from_source_data <- get_number_rows_with_feature_types(vars_unsplit, source_data = filtered_data)
  #Normalize the value per 1000 words
  count_from_source_data <- (count_from_source_data / nrow(filtered_data)) * 1000
  
  log_info("{fn_name}: The count retrieved from the source data is {count_from_source_data}")
  log_info("{fn_name}: count == cell: {signif(count_from_source_data, digits = 2) == signif(cell, digits = 2)}")
}
```

### Ensuring get_var_combos works correctly with different values

## Taking a closer look at the Bellum Hispaniense

I think that there might be some inaccuracies in the parsing on the BH. The data has been trained on a corpus of Classical Latin including Caesar, but nothing resembling the BH. The parser is likely to mistag unusual constructions in the BH incorrectly as a result. Even worse, it might make the text seem more similar to "genuine" Caesar than it actually is.

Let's look at the head of the BH data.

```{r}
View(data |> filter(path == 'spanish'))
```
