---
title: "Postagged Data Accuracy"
bibliography: rmd-notebook.bib
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#101010"
      fg: "#FDF7F7" 
      primary: "#ED79F9"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: scroll
runtime: shiny
params:
  log_threshold: DEBUG
---

```{r setup, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

```{r, load-packages, include=FALSE}
library(shiny)
library(gt)
library(tidyverse)
library(magrittr)
library(rlang)
library(DT)
library(bslib)
library(glue)
library(logger)
conflicted::conflicts_prefer(dplyr::filter, dplyr::lag)

# Set the logger threshold from the yaml header
logger::log_threshold(params$log_threshold)
```

```{r global, include=FALSE}
library(DT) # For rendering my tables

corpus_caesarianum_pathprefix <- "../corpus-caesarianum-authorship/"

# The strings in this look table need to match the names of the two variables defined at the top of the load-data chunk
path_lookup <- c(
  `/home/mdehass/PycharmProjects/corpus-caesarianum-authorship/postagged/postagged-texts.csv`="caesar",
  `/home/mdehass/PycharmProjects/corpus-caesarianum-authorship/postagged/postagged-cicero.csv`="cicero")

caesar <- read_csv(
  glue("{corpus_caesarianum_pathprefix}postagged/postagged-texts.csv")
  )

cicero <- read_csv(
  glue("{corpus_caesarianum_pathprefix}/postagged/postagged-cicero.csv"),
  skip = 1,
  col_names = colnames(caesar)
  )

# We keep the two above tibbles separate so the row numbers still match for Cicero's works
# TODO: Update this code to be susceptible to errors if a column is missing
remove_unused_cols <- function(data) data |> select(-parent_NumType, -parent_Polarity, -parent_Degree, -NumType, -Polarity, -Degree)

caesar <- remove_unused_cols(caesar)
cicero %<>% remove_unused_cols()

col_names_error <- c("book", "row-number", "context", "original_form", "path")
# Add column names from data
col_names_error <- c(col_names_error, colnames(caesar)[
  seq(which(colnames(caesar) == 'form'), which(colnames(caesar) == 'parent_Deprel'))])
col_names_error <- c(col_names_error, "id-correct")

spec <- list(
  book = col_character(),
  `row-number` = col_double(),
  context = col_character(),
  original_form = col_character(),
  path = col_character(),
  .default = col_logical()
)

error_data <- read_csv(
  glue("{corpus_caesarianum_pathprefix}postagged/postag-tests.csv"),
  col_types = spec,
  col_names = col_names_error
  )

# Get rows from original data frames corresponding to the QA'd words in error_data
corresp_error_data <- list_rbind(map2(.x = error_data$`row-number`, .y = path_lookup[error_data$path], .f = function(x, y) env_get(env_parent(), y)[x + 1,]))

# Store options for Data Tables
options <- list(
              scrollX = 300,
              pageHeight = 5
            )

navset_tab(
  nav_panel(title = "Error Data", 
            renderDT(error_data, options = options)),
  nav_panel(title = "Corresponding Original Data",
            renderDT(corresp_error_data, options = options))
)

renderDT(error_data, options = list(
  scrollX = 300,
  pageLength = 5
))
```

# Page 1: Individual Feature Accuracy

## Column

### Accuracy bar plot controls {.tabset data-width="650"}

```{r, percentage-accuracy-per-column}

col_names <- colnames(error_data)[-1:-5]

column_accuracies <- function(error_data) {
  columns_are_logical <- map_lgl(col_names, ~ is_logical(error_data[[.x]]))
  
  if (!all(columns_are_logical)) rlang::abort(message = glue::glue("Error! Not all columns in col_names are logical"))
  
  purrr::map(col_names, .f = ~ sum(na.omit(error_data[[.x]])) / length(na.omit(error_data[[.x]])))
}

to_plot <- data.frame(features = as.factor(col_names), accuracy = unlist(column_accuracies(error_data)))

checkbox_opts <- to_plot$features
names(checkbox_opts) <- checkbox_opts

par(mar=c(1,1,1,1))

splitLayout(
  inputPanel(
    verticalLayout(
      actionButton("select_all", "Select All"),
      actionButton("deselect_all", "Deselect All"),
      checkboxGroupInput(
      "variables",
      "Select variables to view:",
      checkbox_opts,
      selected = checkbox_opts
      )
    ),
    style = "height: 45vh; overflow-y: auto;", 
  )
)

observeEvent(input$select_all, updateCheckboxGroupInput(inputId = "variables", choices = checkbox_opts, selected = checkbox_opts))
  observeEvent(input$deselect_all, updateCheckboxGroupInput(inputId = "variables", choices = checkbox_opts, selected = c()))

```

### Introduction

Given the large amount of automatically annotated data, we cannot review the whole corpus for accuracy. Instead, we have chosen a few hundred words at random and manually reviewed them. With a random sample in hand, we're able to get an approximation of how accurate the data is in the corpus at large. This data allows us to make confident assertions about the accuracy of the data we've used to cluster the texts in the *Corpus Caesarianum*. The morphological data, which includes information like person, number, tense for verbs or case, number, gender for nouns, is on average \_\_\_\_\_% accurate (as seen on page \_\_\_\_\_\_\_), and most errors are located in syntactic structure. Given the role of the syntactic structure in the present analysis, we have increased the accuracy to \_\_\_\_\_% by \_\_\_\_\_\_\_\_\_\_\_. The following dashboard pages present the old accuracy data and new accuracy data after making these automatic adjustments. This transparency allows you, the reader, to see exactly how we've verified the quality of the data before submitting it to the analysis described in our paper.

A summary of the accuracy for each tag individually is given below. The features are those used in the Universal Dependencies CONLLU format [@nivre2020]. The Deprel tag, which is short for *dependency relation* and describes the role in the sentence such as predicate or direct object, is considered accurate if it correctly describes the true relationship in the sentence. The relationship is sometimes correct, though, but with the wrong head. So, a noun can be the subject of a verb and correctly marked *nsubj* in the data, but it is incorrectly considered dependent on another noun instead. GIVE EXAMPLE HERE. If the head word is the correct one, then *parent_form* is marked as correct, whether or not the dependency relation is correct. The graph shows the challenges Stanza has with parsing syntax trees. The morphology on the parent word is highly accurate, but the *Deprel* and *parent_form* can have an accuracy as low as \_\_\_\_\_%.

## Column

### Accuracy bar plot

```{r, barplot}
my_theme <- theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 16), axis.text.y = element_text(size = 12), axis.title = element_text(size = 16))

renderPlot({
    ggplot(to_plot[to_plot$features %in% input$variables,], aes(features, accuracy)) + labs(title = "Percentage accuracy of each feature in the postagged/postag-tests.csv file") + geom_col() + my_theme
  })

```

### 

# Page 2
