---
title: "Postagged Data Accuracy"
bibliography: rmd-notebook.bib
output: 
  flexdashboard::flex_dashboard:
    theme:
      version: 4
      bootswatch: lux
    orientation: columns
    vertical_layout: scroll
runtime: shiny
params:
  log_threshold: DEBUG
---

```{r setup, include=FALSE}
library(flexdashboard)
# Install thematic and un-comment for themed static plots (i.e., ggplot2)
# thematic::thematic_rmd()
```

```{r, load-packages, include=FALSE}
# Install rvest and xml2 if you haven't already
# install.packages("rvest")
# install.packages("xml2")

# Load the packages
library(rvest)
library(xml2)
library(shiny)
library(gt)
library(tidyverse)
library(magrittr)
library(rlang)
library(DT)
library(bslib)
library(glue)
library(logger)
conflicted::conflicts_prefer(dplyr::filter, dplyr::lag)

# Set the logger threshold from the yaml header
logger::log_threshold(params$log_threshold)
```

```{r global, include=FALSE}
library(DT) # For rendering my tables

corpus_caesarianum_pathprefix <- "./" # "../corpus-caesarianum-authorship/"

# The strings in this look table need to match the names of the two variables defined at the top of the load-data chunk
path_lookup <- c(
  `/home/mdehass/PycharmProjects/corpus-caesarianum-authorship/postagged/postagged-texts.csv`="caesar",
  `/home/mdehass/PycharmProjects/corpus-caesarianum-authorship/postagged/postagged-cicero.csv`="cicero",
  `/home/mdehass/PycharmProjects/corpus-caesarianum-authorship/postagged/postagged-sallust.csv`="sallust")

# Lookup table for getting commentary name by author
corpus_lookup <- list(
  `Corpus Caesarianum` = c("gallic", "civil", "african", "alexandrine", "spanish"),
  Cicero = c("amicitia", "senectute", "philippics", "brutus", "deiotaro"),
  Sallust = c("catilinae_coniuratio", "iugurthine")
)

caesar <- read_csv(
  glue("{corpus_caesarianum_pathprefix}postagged/postagged-texts.csv")
  )

cicero <- read_csv(
  glue("{corpus_caesarianum_pathprefix}/postagged/postagged-cicero.csv"),
  skip = 1,
  col_names = colnames(caesar)
  )

sallust <- read_csv(
  glue("{corpus_caesarianum_pathprefix}/postagged/postagged-sallust.csv"),
  skip = 1,
  col_names = colnames(caesar)
)

# We keep the two above tibbles separate so the row numbers still match for Cicero's works
# TODO: Update this code to be susceptible to errors if a column is missing
remove_unused_cols <- function(data) data |> select(-parent_NumType, -parent_Polarity, -parent_Degree, -NumType, -Polarity, -Degree)

caesar <- remove_unused_cols(caesar)
cicero %<>% remove_unused_cols()

col_names_error <- c("book", "row-number", "context", "original_form", "path")
# Add column names from data
col_names_error <- c(col_names_error, colnames(caesar)[
  seq(which(colnames(caesar) == 'form'), which(colnames(caesar) == 'parent_Deprel'))])
col_names_error <- c(col_names_error, "id-correct")

spec <- list(
  book = col_character(),
  `row-number` = col_double(),
  context = col_character(),
  original_form = col_character(),
  path = col_character(),
  .default = col_logical()
)

error_data <- read_csv(
  glue("{corpus_caesarianum_pathprefix}postagged/postag-tests.csv"),
  col_types = spec,
  col_names = col_names_error
  )

# Get rows from original data frames corresponding to the QA'd words in error_data
corresp_error_data <- list_rbind(map2(.x = error_data$`row-number`, .y = path_lookup[error_data$path], .f = function(x, y) env_get(env_parent(), y)[x + 1,]))

# Store options for Data Tables
options <- list(
              scrollX = 300,
              pageHeight = 5
            )

navset_tab(
  nav_panel(title = "Error Data", 
            renderDT(error_data, options = options)),
  nav_panel(title = "Corresponding Original Data",
            renderDT(corresp_error_data, options = options))
)

renderDT(error_data, options = list(
  scrollX = 300,
  pageLength = 5
))
```

# Introduction

## Column

Accuracy for the Stanza model we have used to parse the corpus is freely available online; links and data are available in the table below. Having clear data for the accuracy of our model is crucial. Without confidence in the base data, all of our conclusions rest on a shaky and likely inaccurate foundation. This data shows that the automatically parsed texts have a much higher accuracy with regards to morphology than syntax in the Latin treebanks. Given the role of syntax in our study, we have used this accuracy information and developed our own in order to increase these values.

Stanza evaluates their model by seeing reserving gold standard data from the original, hand-annotated treebanks and seeing whether it gets the same results. To start, here's a brief explanation of each column in the following table:

- **Treebank:** The Universal Dependencies treebank the data is gathered from. The two most relevant are PROIEL and Perseus, which contain extracts from the *Gallic War*, Cicero's *in Catilinam* 1, and Cicero's *Letters to Atticus*.

- **Tokens, Sentences, Words:** The percentage of tokens, sentence, and words that are split up the same way as they are in the original dataset.

- **UPOS, XPOS**: The percentage of POS tags of each type that are the same in the original dataset (XPOS tags are an older format and not used in our data).

- **UFeats:** The percentage of words that have identical features (such as Tense, Voice, Case, Gender, etc.) in the original dataset.

- **AllTags:** The percentage of words that have identical UFeats and \*POS

- **Lemmas:** The percentage of words with the same lemma in the original data (i.e., it identifies the same dictionary entry form in both)

- **UAS:** The unlabeled attachment score, which is the number of words that have the same syntactic parent in the original data.

- **LAS:** The labeled attachment score, which is the number of words that have both the same dependency relation (such as subject or object) and syntactic parent in the original data.

- **CLAS/MLAS/BLEX:** A fuller description can be given on their [website](https://universaldependencies.org/conll18/evaluation.html). Each of these is a more stringent version of the LAS which requires additional features to match (for example, BLEX also requires the lemmas to match).

```{r, include=FALSE}

# Here is the explanation from the script itself:
#   - Tokens: how well do the gold tokens match system tokens
#   - Sentences: how well do the gold sentences match system sentences
#   - Words: how well can the gold words be aligned to system words
#   - UPOS: using aligned words, how well does UPOS match
#   - XPOS: using aligned words, how well does XPOS match
#   - UFeats: using aligned words, how well does universal FEATS match
#   - AllTags: using aligned words, how well does UPOS+XPOS+FEATS match
#   - Lemmas: using aligned words, how well does LEMMA match
#   - UAS: using aligned words, how well does HEAD match
#   - LAS: using aligned words, how well does HEAD+DEPREL(ignoring subtypes) match
#   - CLAS: using aligned words with content DEPREL, how well does
#       HEAD+DEPREL(ignoring subtypes) match
#   - MLAS: using aligned words with content DEPREL, how well does
#       HEAD+DEPREL(ignoring subtypes)+UPOS+UFEATS+FunctionalChildren(DEPREL+UPOS+UFEATS) match
#   - BLEX: using aligned words with content DEPREL, how well does
#       HEAD+DEPREL(ignoring subtypes)+LEMMAS match

stanza_accuracy_data <- "https://stanfordnlp.github.io/stanza/performance.html"

page <- read_html(stanza_accuracy_data)

tables <- page |> html_elements("table") 

tib_stanza_accuracy <- html_table(tables[[1]])

```

```{r, stanza-accuracy-data, collapse=TRUE}

tib_stanza_accuracy |>
  filter(Language == "Latin") |>
  select(-Language, -lcode, -Default) |>
  gt() |>
  tab_header(
    title = "Percentage accuracy per Latin treebank in Stanza's training data", 
    subtitle = md("Scores are defined in the <u>[CoNLL 2018 Shared Task evaluation script](https://universaldependencies.org/conll18/evaluation.html)</u>")) |>
  tab_footnote(footnote = md(glue("Data is available on <u>[{stanza_accuracy_data}]({stanza_accuracy_data})</u>")))

```

We do not have gold standard data to compare against for the overwhelming majority of corpus, so we have approached testing the accuracy a different way. We present a few of the scores in the table below. These are the result of random sampling from the corpus. Words are randomly selected, and each field is marked as correct or incorrect; the random sampling helps ensure that we get a representative collection of words from the corpus and are able to generalize this accuracy information to the whole corpus. The sampling process gets even numbers of words from each text in the corpus (not accounting for books), to ensure that shorter texts like *Bellum Hispaniense* get thoroughly evaluated. We also ensure that each sample from each work represents all the tags in the data. Then, each individual tag is evaluated for accuracy based on comparison with similar structures in the PROIEL treebank or, primarily, the annotation guidelines on universaldependencies.org. If a parent word is not the correct syntactic parent, then I mark the `parent_form` field as incorrect, in case the reader is interested in looking at the data.[^1] Otherwise, the parent fields are marked as correct or incorrect independently of the syntactic child, so those fields will not add any information about the accuracy of the syntactic annotation.

[^1]: `r glue("{corpus_caesarianum_pathprefix}postagged/postag-tests.csv") |> normalizePath()`.

```{r, get-our-corpus-texts}

# Allow user to specify the author; options will be *Corpus Caesarianum*, *Sallust*, and *Cicero*

opts <- unlist(names(corpus_lookup))

# TODO: MAKE THE FOLLOWING CHECKBOXES INSTEAD OF A SELECT BOX, SO YOU CAN PICK MULTIPLE AT ONCE
selectInput(
  "accuracy_table",
  "Select a corpus segment:",
  c("Choose one" = "", opts)
)

render_gt({
  # Get the data; if the default input in the selectizeInput widget is selected, we return the data as-is. Otherwise, the tables get filtered to only the correct titles
  render_error <- if (input$accuracy_table == "") error_data else error_data[corresp_error_data$title %in% corpus_lookup[[input$accuracy_table]],]
  render_corresp <- if (input$accuracy_table == "") corresp_error_data else corresp_error_data[corresp_error_data$title %in% corpus_lookup[[input$accuracy_table]],]
  
  get_vector_accuracy <- function(v) sum(na.omit(v)) / length(na.omit(v))
  # Scores we need to include in our table:
  
  # UPOS: easy enough to get automatically
  
  UPOS_acc <-  get_vector_accuracy(render_error$tag) * 100
  
  # XPOS: NA
  
  # UFeats: easy enough to get automatically
  
  #    Get a function which allows us to purrr::reduce() column-wise until we get a vector the same length as the data frame, but with a TRUE where each row had only TRUEs or NAs and a FALSE otherwise
  reduce_error_data <- function(a, b) {
    # Replace NA with TRUE
    a[is.na(a)] <- TRUE
    b[is.na(b)] <- TRUE
    
    a & b
  }
  
  UFeats_acc <- sum(
    render_error |>
      select(Aspect:Gender) |>
      reduce(reduce_error_data)
  ) / nrow(render_error) * 100
  
  # AllTags: easy enough to get automatically
  AllTags_acc <- sum(
    render_error |>
      select(tag:Gender) |>
      reduce(reduce_error_data)
  ) / nrow(render_error) * 100
  
  # Lemmas: can be done automatically
  Lemmas_acc <-  get_vector_accuracy(render_error$lemma) * 100
  
  # Unlabeled attachment score (UAS): Percentage of words with correct head
  UAS_acc <- get_vector_accuracy(render_error$parent_form) * 100
  
  # Labeled attachment score (LAS): Percentage of words with correct deprel and head after truncation (which is how we account for it too). That is, acl:relcl becomes just acl.
  both_are_accurate <- reduce_error_data(render_error$Deprel, render_error$parent_form)
  LAS_acc <- sum(both_are_accurate) / length(both_are_accurate) * 100
  
  # MLAS adds a few other criteria:
  #   - The UPOS tag is correct
  #   - The following tags are all correct: PronType, NumType, Poss, Reflex, Foreign, Abbr, Gender, Animacy, Number, Case, Definite, Degree, VerbForm, Mood, Tense, Aspect, Voice, Evident, Polarity, Person, Polite
  #   - The words have the same functional children and they both match
  #       - A functional child has one of the following deprels: aux, cop, mark, det, clf, case, cc
  #       - That is, the children have the same: Deprel, UPOS tags, and the features mentioned above
  #   - The accuracy is a percentage out of the total number of "content relations," not just all words. This means that the words in consideration have the following deprels: nsubj, obj, iobj, csubj, ccomp, xcomp, obl, vocative, expl, dislocated, advcl, advmod, discourse, nmod, appos, nummod, acl, amod, conj, fixed, flat, compound, list, parataxis, orphan, goeswith, reparandum, root, dep.
  
  
  # BLEX is like MLAS but the lemmas have to match too
  
  
  # Add to Data Table
  tibble(
    Sample = as.integer(nrow(render_error)),
    UPOS = UPOS_acc,
    UFEATS = UFeats_acc,
    AllTags = AllTags_acc,
    Lemmas = Lemmas_acc,
    UAS = UAS_acc,
    LAS = LAS_acc
  ) |>
    gt() |>
    fmt_number()
})

```

The low LAS accuracy must be improved for a syntactic approach to the Caesar data to be tenable. The remainder of the tabs in this dashboard run through the data to identify the problematic cases, and present the method for correcting them.

# Individual Feature Accuracy {data-orientation="rows"}

## Row

The first step is to discover the situations where the accuracy is low so we can correct them algorithmically or remove them from consideration entirely. That way, we can have data that has a consitency close to the manually annotated source and is unlikely to fit a model to noise. The following graph confirms that the lowest accuracy features are indeed the dependency relation and the head (which shows the percentage of cases where the parent in the syntax tree is the correct one). 

## Row

### Accuracy bar plot controls {.tabset data-width="650"}

```{r, percentage-accuracy-per-column}

col_names <- colnames(error_data)[7:19]
names(col_names) <- col_names
col_names <- c(form = "^form$", col_names, head = "parent_form")


# Create my own version of %in% with partial matching
`%pin%` <- function(x, table) {
  grep(x, table)
}

# Note: not to be used outside of the function below
select_options <- function(label, error_data) {
  columns <- error_data[,label %pin% colnames(error_data)]
  columns |>
    reduce( ~ c(unlist(.x), unlist(.y)))
}

column_accuracies <- function(error_data) {
  all_cols <- map(col_names, ~ select_options(.x, error_data))
  columns_are_logical <- unlist(map(all_cols, ~ is_logical(.x)))
  
  if (!all(columns_are_logical)) rlang::abort(message = glue::glue("Error! Not all columns in col_names are logical"))
  
  purrr::map(col_names, .f = ~ sum(na.omit(select_options(.x, error_data))) / length(na.omit(select_options(.x, error_data))))
}

to_plot <- data.frame(features = as.factor(names(col_names)), accuracy = unlist(column_accuracies(error_data)))

par(mar=c(1,1,1,1))

checkbox_opts <- names(col_names)

splitLayout(
  inputPanel(
    verticalLayout(
      actionButton("select_all", "Select All"),
      actionButton("deselect_all", "Deselect All"),
      checkboxGroupInput(
      "variables",
      "Select variables to view:",
      checkbox_opts,
      selected = checkbox_opts
      )
    ),
    style = "height: 45vh; overflow-y: auto;", 
  )
)

observeEvent(input$select_all, updateCheckboxGroupInput(inputId = "variables", choices = checkbox_opts, selected = checkbox_opts))
  observeEvent(input$deselect_all, updateCheckboxGroupInput(inputId = "variables", choices = checkbox_opts, selected = c()))

```

### Accuracy bar plot

```{r, barplot}
my_theme <- theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 16), axis.text.y = element_text(size = 12), axis.title = element_text(size = 16))


renderPlot({
    ggplot(to_plot[to_plot$features %in% input$variables,], aes(features, accuracy)) + labs(title = "Percentage accuracy of each feature in the postagged/postag-tests.csv file") + geom_col() + my_theme
  })

```



# Examples in context

Fixing the syntactic head with an algorithm is too far beyond the scope of this project. The number of potential heads is vast, and the high level of discontinuity in Latin phrases (well known to any intermediate reader of Horace) means any approach taken here would replace the old errors with new ones. As a result, we should focus on the dependency relation instead. In situations where the head is accurate but the dependency relation isn't, a combination of part of speech and other aspects of the context should help us correct the situation. A noun marked as the head of a complement clause, as shown in the table below, is impossible if it isn't in a phrase with a copula. We can use this information to reduce the number of issues in the corpus.


## Row

```{r, accuracy-display-table}
accuracy <- (nrow(error_data |> filter(Deprel == TRUE, !is.na(Deprel), parent_form == TRUE, !is.na(parent_form))) / nrow(error_data)) * 100

render_gt({
  error_data |>
    mutate(`Dependency Relation` = corresp_error_data$Deprel, head = corresp_error_data$parent_form) |>
    mutate(head = str_sub(head, 8)) |>
    filter(Deprel == FALSE, !is.na(Deprel), parent_form == TRUE, !is.na(parent_form)) |>
    select(context, original_form, `Dependency Relation`, head) |>
    gt() |>
    tab_header(
      title = "Forms with incorrectly parsed dependency relations and correct parent-child links"
      ) 
  
})

  
```

# LAS per feature value

Visualizing the LAS given a specific feature value (such as *Masculine* or *Nominative*) might also help us focus our efforts. The following is a table showing the least accurate features.

```{r, las-per-feature-value}


```


# LAS by Deprel {data-orientation="rows"}

## Row

```{r}
# Add a summary version up here which gives the same chart as below but with all words, not just a specific feature set.

```

## Row

```{r, deprel-accuracy-tests}

# TODO: CHANGE WORDING TO USE 'LAS' INSTEAD OF 'ACCURACY'

# Input: All features except Deprel in a list

# Output: Graph showing accuracy of each Deprel when filtered to only rows which have a value (any value) in the feature selected in Input
features <- colnames(corresp_error_data)

# Two dropdown lists: the first for the feature, and the second populates with values of that feature
inputPanel(
  selectizeInput(
    "feature",
    "Select a feature to filter by:",
    features,
    selected = features[1]
  ),
  selectizeInput(
    "value",
    "Select a value for this feature:",
    c("NA"),
    c("NA")
  )
)

observe({
  # Get the feature from the first dropdown
  x <- input$feature
  
  choices_value <- c('NA', unique(na.omit(corresp_error_data[x])))
  
  updateSelectizeInput(
    inputId = "value",
    choices = choices_value,
    selected = choices_value[1]
  )
  
  
})

renderPlot({
  # Select all valid values for a feature if we pick the default NA value, otherwise retrieve rows where it has a specific feature
  if (input$value == 'NA') {
    row_selection_vector <- !is.na(corresp_error_data[input$feature])
  }  else {
    row_selection_vector <- corresp_error_data[input$feature] == input$value
    row_selection_vector[is.na(row_selection_vector)] <- FALSE
  } 
  
  output_df <- error_data[row_selection_vector,]
  output_corresp <- corresp_error_data[row_selection_vector,]
  output_df$deprel_name <- output_corresp$Deprel
  
  las <- map_dbl(unique(output_df$deprel_name), ~ sum(output_df$Deprel[output_df$deprel_name == .x & output_df$parent_form]) / length(output_df$Deprel[output_df$deprel_name == .x & output_df$parent_form]))
  
  to_graph_df <- tibble(
    deprels = unique(output_df$deprel_name),
    las = las
  )
  
  print("The accuracy values are:\n")
  print(to_graph_df)
  
  # Independent variable is deprel, dependent variable is deprel accuracy 
  ggplot(na.omit(to_graph_df), aes(deprels, las)) +
    labs(
      title = glue("The LAS per dependency relation when it has the {input$feature} feature"),
      subtitle = glue("This feature/value pair occurs {nrow(output_df)} times")) + 
    geom_col() + 
    my_theme
})
```

```{r}
# ADD TABLE WHICH SHOWS THE CONTEXT AND WORD FOR EACH INACCURATE CASE IN THE GRAPH ABOVE
```
